---
title: "Data Analysis with R"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: tango
    fig_caption: true
    df_print: kable
    code_folding: hide
    code_download: true
editor_options: 
  chunk_output_type: console
---

https://www.alliedscientificpro.com/nirvascan
NIRvascan Portable Smart Spectrometer (alliedscientificpro.com)


# Data cream up and form a matrix

## Load Libraries

We load the necessary libraries. If they are not installed, we install them first.

```{r}
if(!require(dplyr)){install.packages('dplyr'); require(dplyr)}
if(!require(tidyr)){install.packages('tidyr'); require(tidyr)}
if(!require(readr)){install.packages('readr'); require(readr)}
if(!require(stringr)){install.packages('stringr'); require(stringr)}
if(!require(tibble)){install.packages('tibble'); require(tibble)}
```

## Set Folder Path

We set the folder path where our data is located.

```{r}
dir.p = 'C:/BSSS_project_lowcost_nir/database'
```

## Import and Merge CSV Files

We import all .csv files from the specified directory and merge them into a single dataframe.

```{r}
getfiles = list.files(path = paste0(dir.p, "/raw_data/spectra_usa_ge/isc/"),
                      pattern="_r.csv", 
                      full.names = T)
```

## Define Function to Combine CSV Files

We define a function that reads each CSV file and combines the data into a single dataframe.

```{r}
combineCSVFiles = function(filePaths){
  combinedData = data.frame() # empty dataframe
  
  for (filePath in filePaths){
    
    data = read_csv(filePath, id = "path", skip = 28) # read each CSV file
    
    combinedData = rbind(combinedData, data) # Combine the data
  }
  return(combinedData)
}
```

## Use Function to Combine CSV Files

We use the function to combine the CSV files into a single dataframe.

```{r}
combinedDF = combineCSVFiles(getfiles)
```

## Check the Final Dataframe

We print and view the final dataframe.

```{r}
print(combinedDF)
View(combinedDF)
```

## Create ID Column

We create an ID column using the named ID during recording spec.

```{r}
combinedDF <- combinedDF %>%
  dplyr::mutate(id = str_extract(sub("\\..*$", "", basename(combinedDF$path)),
                                 "(?<=_)(\\d+)(?=\\_)"))

combinedDF$id = paste("isc", combinedDF$id, sep = "_") # Check if it is necessary or not

combinedDF = combinedDF %>%
  dplyr::select(c('id'), everything())

combinedDF.ok = combinedDF[, -2]
colnames(combinedDF.ok)[2:3] = c("wavelength", "reflectance")

View(combinedDF.ok)
combinedDF.ok[combinedDF.ok$id == c("00000", "00180"),]
```

## Create a New Column

We create a new column by extracting the numbers after the last comma.

```{r}
df <- combinedDF.ok %>%
  mutate(NewColumn = paste0(",", sub(".+,(\\d+)$", "\\1", `reflectance`)))

df$NewColumn = as.numeric(gsub(",", ".", df$NewColumn))

df.ok = df[, -3]
df.ok[1:10,]

colnames(df.ok)[3] = "reflectance"
```

## Reshape Dataframe

We reshape the dataframe from long to wide format.

```{r}
combined.df.ok = df.ok

colnames(combined.df.ok)
combinedDF.wide = tidyr::spread(combined.df.ok,
                                key = "id",
                                value = "reflectance")

row_names <- colnames(combinedDF.wide)[-1]

transposed_tibble <- as_tibble(t(as.matrix(combinedDF.wide)))

colnames(transposed_tibble) = transposed_tibble[1,]
transposed_tibble[2:5, c(1:4, ncol(transposed_tibble))]

combined.df.org = transposed_tibble[-1, ]
combined.df.org[1:5, c(1:5, ncol(combined.df.org))]
combined.df.org$id = row_names

combined.df.org = combined.df.org %>%
  select(id, everything())

combined.df.org[1:5, c(1:5, ncol(combined.df.org))]
```

## Export Dataframe

We export the final dataframe to a CSV file.

```{r}
write.csv(combined.df.org, paste0(dir.p, "/raw_data/spectra_usa_ge/isc/combinedSpec.csv"))
```

## Save R Environment

We save the R environment for future use.

```{r}
save.image(paste0(dir.p, "/r_code/isc_importing_and_cleaning.RData"))
```
