---
title: "Data Analysis with R"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: tango
    fig_caption: true
    df_print: kable
    code_folding: hide
    code_download: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries

We load the necessary libraries. If they are not installed, we install them first.

```{r}
if(!require(dplyr)){install.packages('dplyr'); require(dplyr)}
if(!require(tidyr)){install.packages('tidyr'); require(tidyr)}
if(!require(prospectr)){install.packages('prospectr'); require(prospectr)}
if(!require(pls)){install.packages('pls'); require(pls)}
if(!require(resemble)){install.packages('resemble'); require(resemble)}
if(!require(clhs)){install.packages('clhs'); require(clhs)}
if(!require(ggplot2)){install.packages('ggplot2'); require(ggplot2)}
```

### Set Folder Path

We set the folder path where our data is located.

```{r}
dir.p = 'D:/project_lowcost_nir/database' # homeoffice/
```

### Importing data from ASD spectrometer

We import the data from the ASD spectrometer and reorganise the dataframe.

```{r}
asd.osc = read.delim(paste0(dir.p, "/raw_data/oscar/asd/osc_asd.txt"), header = T, sep = ";")
head(asd.osc[, 1:10], 5)

asd.osc.df = as.data.frame(t(asd.osc))
colnames(asd.osc.df) = as.integer(asd.osc.df[1, ])
asd.osc.df = asd.osc.df[-1,]
asd.osc.df$id = rownames(asd.osc.df)

asd.osc.df = asd.osc.df %>%
  dplyr::select(c('id'), everything())

asd.osc.df$id = stringr::str_extract(asd.osc.df$id, "(?<=_)(\\d+)(?=\\_)")

asd.osc.df$id = paste("osc", asd.osc.df$id, sep = "_")

asd.osc.df[1:5, c(1:5, ncol(asd.osc.df))]

asd.osc.vnir = aggregate(.~id, data = asd.osc.df, FUN = mean) # averaging the 3 repetitions

asd.osc.vnir[1:5, c(1:5, ncol(asd.osc.vnir))]

# write.csv(asd.osc.vnir, paste0(dir.p, '/tables/asd_vnir_oscar_100pts.csv'))
```

### Importing data from ISC NIRScan spectrometer

We import the data from the ISC NIRScan spectrometer and reorganise the dataframe.

```{r}
isc.osc = read.csv(paste0(dir.p, "/raw_data/oscar/isc/combinedSpec.csv"), 
                   header = T, sep = ",")[-1]

head(isc.osc[, 1:10], 5)

isc.osc.df = as.data.frame(t(isc.osc))
colnames(isc.osc.df) = as.integer(isc.osc.df[1, ])
isc.osc.df = isc.osc.df[-1,]

isc.osc.df$id = rownames(isc.osc.df)

isc.osc.df = isc.osc.df %>%
  dplyr::select(c('id'), everything())

isc.osc.df[1:5, c(1:5, ncol(isc.osc.df))]

isc.osc.nir = isc.osc.df

rownames(isc.osc.nir) = NULL

isc.osc.nir[1:5, c(1:5, ncol(isc.osc.nir))]

# write.csv(isc.osc.nir, paste0(dir.p, '/tables/isc_nir_oscar_100pts.csv'))
```

#### Plotting both spectra

We plot the spectra for both the ASD and ISC NIRScan spectrometers.

```{r}
asd.osc.vnir[1:5, c(1:5, ncol(asd.osc.vnir))]

# jpeg(paste0(dir.p, '/results/figs/original_asd_osc_vnir.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(asd.osc.vnir)[2:ncol(asd.osc.vnir)]), 
        t(asd.osc.vnir[,2:ncol(asd.osc.vnir)]),
        xlim = c(350, 2500),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)

axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)
mtext('Original ASD Spectra')
grid(lwd=1, nx=50, ny=50)
# dev.off()
```

```{r}
isc.osc.nir[1:5, c(1:5, ncol(isc.osc.nir))]

# jpeg(paste0(dir.p, '/results/figs/original_isc_osc_nir.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(isc.osc.nir)[2:ncol(isc.osc.nir)]), 
        t(isc.osc.nir[,2:ncol(isc.osc.nir)]),
        xlim = c(350, 2500),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.0, green = 0.4, blue = 0.6, alpha = 0.4),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2,
        add = T)
axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)
mtext('Original NIRScan Spectra')
grid(lwd=1, nx=50, ny=50)
# dev.off()
```

#### All together

```{r}
# jpeg(paste0(dir.p, '/results/figs/original_asd_vnir_isc_nir_osc.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(asd.osc.vnir)[2:ncol(asd.osc.vnir)]), 
        t(asd.osc.vnir[1:3,2:ncol(asd.osc.vnir)]),
        #xlim = c(400, 2450),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = c("red", "blue", "grey"), #rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'Reflectance Factor (x100)',
        las=1, font=2)

axis(1, at = seq(400, 2450, by = 50), labels = T, las = 2)

matplot(as.numeric(colnames(isc.osc.nir)[2:ncol(isc.osc.nir)]), 
        t(isc.osc.nir[1:3,2:ncol(isc.osc.nir)]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 2,
        col = c("red", "blue", "grey"), #rgb(red = 0.0, green = 0.4, blue = 0.6, alpha = 0.4),
        las=1, font=2,
        add = T)
legend("topright", legend = c("Sample 1", "Sample 2", "Sample 3"), col = c("red", "blue", "grey"), lty = 1, bty = "n")
# dev.off()
```

### Pre-processing the spectra ----

#### Splice Correction applicable for ASD only

```{r}
asd.osc.vnir[1:5, c(1:5, ncol(asd.osc.vnir))]

db.spl.asd.osc = asd.osc.vnir
colnames(db.spl.asd.osc)[c(1:4, ncol(db.spl.asd.osc))]
rownames(db.spl.asd.osc) = db.spl.asd.osc$id

head(db.spl.asd.osc[, c(1:4, ncol(db.spl.asd.osc))])

class(db.spl.asd.osc)
colnames(db.spl.asd.osc)[1:10]
wav.spl.asd = as.numeric(colnames(db.spl.asd.osc[, 2:ncol(db.spl.asd.osc)]))
db.spl.asd.osc.m = as.matrix(db.spl.asd.osc[, 2:ncol(db.spl.asd.osc)]) # assembling matrix
class(db.spl.asd.osc.m) # checking class

db.spl.asd.osc.c = prospectr::spliceCorrection(db.spl.asd.osc.m, wav.spl.asd, 
                                               splice = c(1000, 1800), interpol.bands = 10)
class(db.spl.asd.osc.c)

db.spl.asd.osc.d = as.data.frame(db.spl.asd.osc.c) # converting back to DF
db.spl.asd.osc.d[1:20, 1:5]
db.spl.asd.osc.d$id = rownames(db.spl.asd.osc.d)
db.spl.asd.osc.d = db.spl.asd.osc.d %>%
  dplyr::select(c('id'), everything())

colnames(db.spl.asd.osc.d)[52:53]

db.spl.asd.osc.d[1:5, c(1:3, 52:53, ncol(db.spl.asd.osc.d))]

jpeg(paste0(dir.p, '/results/figs/splice_asd_osc_vnir.jpeg'), 
     width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(db.spl.asd.osc.d)[2:ncol(db.spl.asd.osc.d)]), 
        t(db.spl.asd.osc.d[,2:ncol(db.spl.asd.osc.d)]),
        xlim = c(350, 2500),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)
axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)
mtext('Splice Correction ASD Spectra')
grid(lwd=1, nx=50, ny=50)
dev.off()

# write.csv(db.spl.asd.osc.d, paste0(dir.p, '/tables/asd_vnir_splicecorrection_oscar_100pts.csv'))
```

### Savitzky-Golay filtering algorithm

[A **Savitzky--Golay filter** is a digital filter that can be applied to a set of digital data points for the purpose of smoothing the data, that is, to increase the precision of the data without distorting the signal tendency^1^](https://en.wikipedia.org/wiki/Savitzky%E2%80%93Golay_filter). [This is achieved, in a process known as convolution, by fitting successive sub-sets of adjacent data points with a low-degree polynomial by the method of linear least squares^1^](https://en.wikipedia.org/wiki/Savitzky%E2%80%93Golay_filter).

[When the data points are equally spaced, an analytical solution to the least-squares equations can be found, in the form of a single set of "convolution coefficients" that can be applied to all data sub-sets, to give estimates of the smoothed signal, (or derivatives of the smoothed signal) at the central point of each sub-set^1^](https://en.wikipedia.org/wiki/Savitzky%E2%80%93Golay_filter).

The method, based on established mathematical procedures, was popularized by Abraham Savitzky and Marcel J. E. [Golay, who published tables of convolution coefficients for various polynomials and sub-set sizes in 1964^1^](https://en.wikipedia.org/wiki/Savitzky%E2%80%93Golay_filter). [The method has been extended for the treatment of 2- and 3-dimensional data^1^](https://en.wikipedia.org/wiki/Savitzky%E2%80%93Golay_filter).

Savitzky-Golay filters perform better in some applications than standard averaging FIR filters, which tend to filter high-frequency content along with the noise. [Savitzky-Golay filters are more effective at preserving high frequency signal components but less successful at rejecting noise^2^](https://www.mathworks.com/help/signal/ref/sgolayfilt.html).

#### (1) ASD spectrometer

```{r}
db.spl.asd.osc.d[1:5, c(1:5, ncol(db.spl.asd.osc.d))]
colnames(db.spl.asd.osc.d[, 1:20])

db.asd.osc.sg = prospectr::savitzkyGolay(db.spl.asd.osc.d[, 2:ncol(db.spl.asd.osc.d)],
                                         m = 0, # differentiation order (derivative)
                                         p = 2, # polynomial order
                                         w = 11 # window size
                                         ) 
class(db.asd.osc.sg)
db.asd.osc.sg = as.data.frame(db.asd.osc.sg)
db.asd.osc.sg[1:5, c(1:5, ncol(db.asd.osc.sg))]
db.asd.osc.sg$id = rownames(db.asd.osc.sg)
head(db.asd.osc.sg[, c(1:5, ncol(db.asd.osc.sg))])
#colnames(db.train.sg) = gsub("X", "", names(db.train.sg))

db.asd.osc.sg = db.asd.osc.sg %>%
  select(c("id"), everything())

head(db.asd.osc.sg[, c(1:6, ncol(db.asd.osc.sg))])

jpeg(paste0(dir.p, '/results/figs/sg_2ord_11pts_0der_asd_osc_vnir.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(db.asd.osc.sg)[2:ncol(db.asd.osc.sg)]), 
        t(db.asd.osc.sg[,2:ncol(db.asd.osc.sg)]),
        xlim = c(350, 2500),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)
axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)
mtext('ASD Spectra: Savitzky-Golay 2nd order polynomial and window size of 11 smoothing points')
grid(lwd=1, nx=50, ny=50)
dev.off()

write.csv(db.asd.osc.sg, paste0(dir.p, '/tables/asd_vnir_sg_2p_w11_0d_oscar_100pts.csv'))
```

#### (2) NIRScan spectrometer

```{R}
isc.osc.nir[1:5, c(1:5, ncol(isc.osc.nir))]
colnames(isc.osc.nir[, 1:20])

rownames(isc.osc.nir) = isc.osc.nir$id

db.isc.osc.sg = prospectr::savitzkyGolay(isc.osc.nir[, 2:ncol(isc.osc.nir)],
                                         m = 0, # differentiation order (derivative)
                                         p = 2, # polynomial order
                                         w = 11 # window size
) 
class(db.isc.osc.sg)
db.isc.osc.sg = as.data.frame(db.isc.osc.sg)
db.isc.osc.sg[1:5, c(1:5, ncol(db.isc.osc.sg))]
db.isc.osc.sg$id = rownames(db.isc.osc.sg)
head(db.isc.osc.sg[, c(1:5, ncol(db.isc.osc.sg))])
#colnames(db.train.sg) = gsub("X", "", names(db.train.sg))

db.isc.osc.sg = db.isc.osc.sg %>%
  select(c("id"), everything())

head(db.isc.osc.sg[, c(1:6, ncol(db.isc.osc.sg))])

# jpeg(paste0(dir.p, '/results/figs/sg_2ord_11pts_0der_isc_osc_vnir.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(db.isc.osc.sg)[2:ncol(db.isc.osc.sg)]), 
        t(db.isc.osc.sg[,2:ncol(db.isc.osc.sg)]),
        xlim = c(350, 2500),
        #ylim = c(0.0, 0.7),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)
axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)
mtext('NIRScan Spectra: Savitzky-Golay 2nd order polynomial and window size of 11 smoothing points')
grid(lwd=1, nx=50, ny=50)
# dev.off()

# write.csv(db.isc.osc.sg, paste0(dir.p, '/tables/isc_vnir_sg_2p_w11_0d_oscar_100pts.csv'))
```

#### (3) All in one - Plotting

```{R}
# jpeg(paste0(dir.p, '/results/figs/sg_asd_vnir_isc_nir_osc.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(db.asd.osc.sg)[2:ncol(db.asd.osc.sg)]), 
        t(db.asd.osc.sg[1:3,2:ncol(db.asd.osc.sg)]),
        #xlim = c(400, 2450),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = c("red", "blue", "grey"), #rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3), 
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'Reflectance Factor (x100)',
        las=1, font=2)
axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)

matplot(as.numeric(colnames(db.isc.osc.sg)[2:ncol(db.isc.osc.sg)]), 
        t(db.isc.osc.sg[1:3,2:ncol(db.isc.osc.sg)]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 2,
        col = c("red", "blue", "grey"), #rgb(red = 0.0, green = 0.4, blue = 0.6, alpha = 0.4),
        las=1, font=2,
        add = T)
legend("topright", legend = c("Sample 1", "Sample 2", "Sample 3"), col = c("red", "blue", "grey"), lty = 1, bty = "n")
# dev.off()
```

### Selecting the same spectral range (950 - 1650 nm) ----

In spectroscopy, it's vital to consistently capture the same spectral range (950 - 1650 nm) across different devices for accurate data comparison and modelling. This involves calibrating each device using a known light source, adjusting settings or filters to capture the desired range, and processing the data to account for device-specific sensitivities. Despite capturing the same range, inherent device characteristics may cause differences, which should be considered in data comparisons or models.

```{R}
## ASD
db.asd.osc.sg[1:3, c(596:597, 1296:1297)]
db.asd.osc.ok = db.asd.osc.sg[, 597:1297]
db.asd.osc.ok[1:3, c(1:4, ncol(db.asd.osc.ok))]

## NIRScan
db.isc.osc.sg[1:3, c(9:10, 209:211)]
db.isc.osc.ok = db.isc.osc.sg[, 9:209]
db.isc.osc.ok[1:3, c(1:4, ncol(db.isc.osc.ok))]
```

### Resampling them to the same window size of 5 nm ----

Resampling the spectral data to the same window size of 5 nm is a common practice in spectroscopy to ensure consistency across different devices. This process involves interpolating the original data points to generate new data points at regular intervals of 5 nm. This can help to align the spectral data from different devices, making it easier to compare and model the data. However, it's important to note that resampling can sometimes lead to a loss of information, especially if the original data was captured at a higher resolution. Therefore, the choice of window size should be carefully considered based on the specific requirements of your analysis.

#### ASD from 1 nm

```{R}

wav.asd = as.numeric(colnames(db.asd.osc.ok)) # retrieving wavelengths

db.asd.osc.r = as.matrix(db.asd.osc.ok)

#rownames(db.asd.osc.r) = db.asd.osc.ok$id

db.asd.osc.r = prospectr::resample(db.asd.osc.r, 
                                   wav.asd, 
                                   seq(950, 1650, 5), 
                                   interpol = "spline") # window size of 5 nm
class(db.asd.osc.r)
db.asd.osc.r = as.data.frame(db.asd.osc.r)

db.asd.osc.r$id = rownames(db.asd.osc.r)
db.asd.osc.r = db.asd.osc.r %>%
  select(c("id"), everything())

db.asd.osc.r[1:5, c(1:5, ncol(db.asd.osc.r))]
```

#### NIRScan from 3 nm

```{R}
wav.isc = as.numeric(colnames(db.isc.osc.ok)) # retrieving wavelengths

db.isc.osc.r = as.matrix(db.isc.osc.ok)

#rownames(db.isc.osc.r) = db.isc.osc.ok$id

db.isc.osc.r = prospectr::resample(db.isc.osc.r, 
                                   wav.isc, 
                                   seq(950, 1650, 5), 
                                   interpol = "spline") # window size of 5 nm
class(db.isc.osc.r)
db.isc.osc.r = as.data.frame(db.isc.osc.r)

db.isc.osc.r$id = rownames(db.isc.osc.r)
db.isc.osc.r = db.isc.osc.r %>%
  select(c("id"), everything())

db.isc.osc.r[1:5, c(1:5, ncol(db.isc.osc.r))]
```

#### Plot all together - plotting

```{R}
# jpeg(paste0(dir.p, '/results/figs/resampled_asd_isc_nir_osc.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(db.asd.osc.r)[2:ncol(db.asd.osc.r)]), 
        t(db.asd.osc.r[,2:ncol(db.asd.osc.r)]),
        #xlim = c(400, 2450),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3), #c("red", "blue", "grey"),
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'Reflectance Factor (x100)',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(as.numeric(colnames(db.isc.osc.r)[2:ncol(db.isc.osc.r)]), 
        t(db.isc.osc.r[,2:ncol(db.isc.osc.r)]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 2,
        col = rgb(red = 0.6, green = 0.4, blue = 0.2, alpha = 0.4), #c("red", "blue", "grey"), 
        las=1, font=2,
        add = F)
legend("topright", legend = c("ASD", "NIRScan"), 
       col = c(rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
               rgb(red = 0.6, green = 0.4, blue = 0.2, alpha = 0.4)), 
       lty = 1, bty = "n")
# dev.off()
```

### Splitting the dataset using Conditional Latin Hypercube Sampling (cLHS) ----

[Conditional Latin Hypercube Sampling (cLHS) is a method used to split datasets in a way that ensures a representative sample for model calibration and validation^1^](https://www.mdpi.com/2072-4292/15/11/2915). [This method is particularly useful when dealing with multivariate data, as it helps to maintain the multivariate structure of the data in the subsets^1^](https://www.mdpi.com/2072-4292/15/11/2915). [In practice, cLHS can be used to split a dataset into two subsets, for example, 80% for model calibration (or training) and 20% for internal validation^1^](https://www.mdpi.com/2072-4292/15/11/2915). [This allows for the development of robust models that are well-calibrated and validated on independent data^1^](https://www.mdpi.com/2072-4292/15/11/2915).

It\'s important to note that the percentages used for splitting the data can be adjusted based on the specific requirements of your analysis. [The key is to ensure that both subsets of data are representative of the overall dataset, allowing for accurate modelling and comparison^1^](https://www.mdpi.com/2072-4292/15/11/2915).

```{R}
db.asd.osc.r
colnames(db.asd.osc.r)[1:20]
db.asd.osc.r[1:5, 1:5]

#### Performing PC analysis before splitting the data into cal and val using ASD as standard

maxexplvar = 0.999 # indicating the max amount of cumulative variance explained
asd.osc.pc.spec = resemble::pc_projection(Xr = as.matrix(db.asd.osc.r[, -1]),
                                                 pc_selection = list("cumvar", maxexplvar),
                                                 center = T,
                                                 scale = F)


### Conditioned Latin Hypercube Sampling (cLHS)

nrow(db.asd.osc.r) # 100 obs.
class(asd.osc.pc.spec$scores)

# Soil attributes with 262 obs.
asd.osc.clhs = clhs::clhs(x = as.data.frame(asd.osc.pc.spec$scores),
                          size = round(0.70*nrow(db.asd.osc.r)), 
                          iter = 1000, simple = FALSE)

# jpeg(paste0(dir.p, '/results/figs/asd_osc_clhs.jpeg'), width = 10, height = 10, units = 'in' , res = 300)

## plot the 2 PCs
# Create a data frame for the first two PCs and calibration samples
pc_scores = as.data.frame(asd.osc.pc.spec$scores[, 1:2])
calibration_samples = as.data.frame(asd.osc.pc.spec$scores[, 1:2][asd.osc.clhs$index_samples, ])

# Add explained variances and cumulative explained variances to the plot
explained_var = as.data.frame(asd.osc.pc.spec$variance$x_var[2, ]) #explained_var
cumulative_explained_var = as.data.frame(asd.osc.pc.spec$variance$x_var[3, ]) #cumulative_explained_var

# Plot the first two PCs
ggplot(pc_scores, aes(pc_1, pc_2)) +
  geom_point(color = "gray", alpha = 0.5, size = 3) +
  geom_point(data = calibration_samples, color = "blue", size = 3) +
  theme_minimal() +
  labs(x = "PC 1 (99.5%)", y = "PC 2 (0.40%)") +
  ggtitle("Scatter Plot of PC 1 and PC 2") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") #+
  #geom_text(aes(label = paste0("Explained variance: ", round(explained_var, 3))), 
  #          x = max(pc_scores$pc_1), y = max(pc_scores$pc_2), hjust = 1, vjust = 1) +
  #geom_text(aes(label = paste0("Cumulative explained variance: ", round(cumulative_explained_var, 3))), 
  #          x = max(pc_scores$pc_1), y = max(pc_scores$pc_2) - 0.01, hjust = 1, vjust = 1)

# dev.off()
```

### ASD calibration

```{R}
db.asd.osc.r.cal = db.asd.osc.r[asd.osc.clhs$index_samples, ] 
nrow(db.asd.osc.r.cal) # 70% / 70
# write.csv(db.asd.osc.r.cal, paste0(dir.p, '/tables/asd_nir_sg_cal_oscar_70pts.csv'))

db.asd.osc.sg.cal = db.asd.osc.sg[db.asd.osc.sg$id %in% db.asd.osc.r.cal$id,]
# write.csv(db.asd.osc.sg.cal, paste0(dir.p, '/tables/db_model/asd_visnir_sg_cal_oscar_70pts.csv'))

db.asd.osc.r.val = db.asd.osc.r[-asd.osc.clhs$index_samples, ] 
nrow(db.asd.osc.r.val) # 30% / 30
# write.csv(db.asd.osc.r.val, paste0(dir.p, '/tables/asd_nir_sg_cal_oscar_30pts.csv'))

db.asd.osc.sg.val = db.asd.osc.sg[db.asd.osc.sg$id %in% db.asd.osc.r.val$id,]
# write.csv(db.asd.osc.sg.val, paste0(dir.p, '/tables/db_model/asd_visnir_sg_val_oscar_70pts.csv'))


# Selecting the same samples from ISC NIRScan
db.isc.osc.r
colnames(db.isc.osc.r)
colnames(db.asd.osc.r.cal)
db.isc.osc.r.cal = db.isc.osc.r[db.isc.osc.r$id %in% db.asd.osc.r.cal$id,]
write.csv(db.isc.osc.r.cal, paste0(dir.p, '/tables/isc_nir_sg_cal_oscar_70pts.csv'))

db.isc.osc.r.val = db.isc.osc.r[db.isc.osc.r$id %in% db.asd.osc.r.val$id,]
# write.csv(db.isc.osc.r.val, paste0(dir.p, '/tables/isc_nir_sg_val_oscar_30pts.csv'))
```

#### Organising the data into source and target calibration and validation and performing SNV ----

### CALIBRATION

```{R}
### SOURCE
source.spectra = data.frame(db.isc.osc.r.cal[, 1], NIR = I(db.isc.osc.r.cal[,-1]))
colnames(source.spectra)[1] = "id"
write.csv(source.spectra, paste0(dir.p, '/tables/db_model/db_source_isc_spec_cal.csv'))
#### SNV
source.snv = source.spectra
source.snv$NIR = prospectr::standardNormalVariate(source.snv$NIR)
colnames(source.snv)
write.csv(source.snv, paste0(dir.p, '/tables/db_model/db_source_isc_spec_snv_cal.csv'))

### TARGET
target.spectra = data.frame(db.asd.osc.r.cal[, 1], NIR = I(db.asd.osc.r.cal[,-1]))
colnames(target.spectra)[1] = "id"
write.csv(target.spectra, paste0(dir.p, '/tables/db_model/db_target_asd_spec_cal.csv'))
#### SNV
target.snv = target.spectra
target.snv$NIR = prospectr::standardNormalVariate(target.snv$NIR)
colnames(target.snv)
write.csv(target.snv, paste0(dir.p, '/tables/db_model/db_target_asd_spec_snv_cal.csv'))

wavelength = as.numeric(substring(colnames(source.snv$NIR), 1, 7))
str(source.snv)
class(source.snv)
```

### VALIDATION

```{R}
### SOURCE
source.spectra.val = data.frame(db.isc.osc.r.val[, 1], NIR = I(db.isc.osc.r.val[,-1]))
colnames(source.spectra.val)[1] = "id"
write.csv(source.spectra.val, paste0(dir.p, '/tables/db_model/db_source_isc_spec_val.csv'))
### SNV
source.snv.val = source.spectra.val
source.snv.val$NIR = prospectr::standardNormalVariate(source.snv.val$NIR)
colnames(source.snv.val)
write.csv(source.snv.val, paste0(dir.p, '/tables/db_model/db_source_isc_spec_snv_val.csv'))

### TARGET
target.spectra.val = data.frame(db.asd.osc.r.val[, 1], NIR = I(db.asd.osc.r.val[,-1]))
colnames(target.spectra.val)[1] = "id"
write.csv(target.spectra.val, paste0(dir.p, '/tables/db_model/db_target_asd_spec_val.csv'))

#### SNV
target.snv.val = target.spectra.val
target.snv.val$NIR = prospectr::standardNormalVariate(target.snv.val$NIR)
colnames(target.snv.val)
write.csv(target.snv.val, paste0(dir.p, '/tables/db_model/db_target_asd_spec_snv_val.csv'))
```

### Substraction transfer, spectral outliers removed by IQR -----

In spectroscopy, Subtraction Transfer is used to isolate the spectral contribution of a specific component, such as water at the 1415 nm band, from a mixture. However, spectral data often contains outliers that can distort the results. To address this, the Interquartile Range (IQR) method is used to identify and remove outliers.

The IQR method calculates the first and third quartiles of the data, and any data point outside 1.5 times the IQR is considered an outlier and removed. This results in a clean, outlier-free spectrum for further analysis and modeling. However, care should be taken as inappropriate removal of outliers can lead to loss of information.

```{R}

common.ids = intersect(db.asd.osc.r.cal$id, db.isc.osc.r.cal$id)

## Through substraction
subtracted.spectra.asd.isc = data.frame(id = common.ids)

for (col in names(target.spectra$NIR)) {  # Exclude the ID column
  subtracted.spectra.asd.isc[[col]] = round(target.spectra$NIR[target.spectra$id %in% common.ids, col] - 
                                      source.spectra$NIR[source.spectra$id %in% common.ids, col],3)
}

# Print the subtracted dataframe
print(subtracted.spectra.asd.isc)

summary(subtracted.spectra.asd.isc[, c("1400", "1415")])
boxplot(subtracted.spectra.asd.isc[, "1415"])

# Removing outliers using as reference water band 1415 nm
quartiles = quantile(subtracted.spectra.asd.isc[, "1415"], probs=c(.25, .75), rm.na=T)
iqr = IQR(subtracted.spectra.asd.isc[, "1415"])
lv = quartiles[1]-1.5*iqr
uv = quartiles[2]+1.5*iqr

subtracted.spectra.asd.isc.noout = subset(subtracted.spectra.asd.isc, subtracted.spectra.asd.isc[, "1415"] > lv & subtracted.spectra.asd.isc[, "1415"] < uv)
boxplot(subtracted.spectra.asd.isc.noout[, "1415"])
hist(subtracted.spectra.asd.isc.noout[, "1415"])
summary(subtracted.spectra.asd.isc.noout[, c("1400", "1415")])
isc.subtracted.median = apply(subtracted.spectra.asd.isc.noout[-1], 2, median)
isc.factor = isc.subtracted.median['1415'] # Water band 1400 nm or *1415nm

source.spectra.factor.cor = source.spectra

source.spectra.factor.cor$NIR = source.spectra.factor.cor$NIR + isc.factor
```

#### All in one Calibration

```{r}
jpeg(paste0(dir.p, '/results/figs/MEAN_SPEC_SIGN_factor_correction_nir_asd_isc_osc_calibration.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(wavelength, 
        colMeans(target.spectra$NIR),
        #xlim = c(400, 2450),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 3,
        lwd = 2,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.8),
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'Reflectance Factor (x100)',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(wavelength, 
        colMeans(source.spectra$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)

matplot(wavelength, 
        colMeans(source.spectra.factor.cor$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 1.0, green = 0.0, blue = 0.0, alpha = 0.5),
        las=1, font=2,
        add = T)

legend("topleft",bty="n", legend=c("Target: ASD data","Source: ISC NIRScan data", "Index: Fitting source in target data"),
       col=c(rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
             "forestgreen",
             "red"),
       lty=c(1),cex=1,lwd=3)

dev.off()
```

#### VALIDATION

```{R}
source.spectra.factor.cor.val = source.spectra.val

source.spectra.factor.cor.val$NIR = source.spectra.factor.cor.val$NIR + isc.factor
```

#### All in one Validation

```{r}
jpeg(paste0(dir.p, '/results/figs/MEAN_SPEC_SIGN_factor_correction_nir_asd_isc_osc_validation.jpeg'), width = 10, height = 10, units = 'in', res = 300)

matplot(wavelength, 
        colMeans(target.spectra.val$NIR),
        #xlim = c(400, 2450),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 3,
        lwd = 2,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.8),
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'Reflectance Factor (x100)',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(wavelength, 
        colMeans(source.spectra.val$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)

matplot(wavelength, 
        colMeans(source.spectra.factor.cor.val$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 1.0, green = 0.0, blue = 0.0, alpha = 0.5),
        las=1, font=2,
        add = T)

legend("topleft",bty="n", legend=c("Target: ASD data",
                                   "Source: ISC NIRScan data", 
                                   "Index: Fitting source in target data"),
       col=c(rgb(red = 0.5, green = 0.5, 
                 blue = 0.5, alpha = 0.3),
             "forestgreen","red"),
       lty=c(1),cex=1,lwd=3)

dev.off()
```

```{R}
### Exporting data.frame
source.spectra.factor.cor
write.csv(source.spectra.factor.cor, paste0(dir.p, '/tables/db_model/db_source_spec_subcorr_cal.csv'))

source.spectra.factor.cor.val
write.csv(source.spectra.factor.cor.val, paste0(dir.p, '/tables/db_model/db_source_spec_subcorr_val.csv'))

## SG-SNV CALIBRATION
subtracted.snv.asd.isc = data.frame(id = common.ids)

#colnames(target.snv$NIR)
for (col in colnames(target.snv$NIR)) {  # Exclude the ID column
  subtracted.snv.asd.isc[[col]] = round(target.snv$NIR[target.snv$id %in% common.ids, col] - 
                                              source.snv$NIR[source.snv$id %in% common.ids, col],3)
}

# Print the subtracted dataframe
print(subtracted.snv.asd.isc)

summary(subtracted.snv.asd.isc[, c("1400", "1415")])
boxplot(subtracted.snv.asd.isc[, "1415"])
```

### Removing outliers using as reference water band 1415 nm

```{r}
quartiles.snv = quantile(subtracted.snv.asd.isc[, "1415"], probs=c(.25, .75), rm.na=T)
iqr.snv = IQR(subtracted.snv.asd.isc[, "1415"])
lv.snv = quartiles[1]-1.5*iqr
uv.snv = quartiles[2]+1.5*iqr

subtracted.snv.asd.isc.noout = subset(subtracted.snv.asd.isc, subtracted.snv.asd.isc[, "1415"] > lv.snv & subtracted.snv.asd.isc[, "1415"] < uv.snv)
boxplot(subtracted.snv.asd.isc.noout[, "1415"])
hist(subtracted.snv.asd.isc.noout[, "1415"])
summary(subtracted.snv.asd.isc.noout[, c("1400", "1415")])
isc.subtracted.snv.median = apply(subtracted.snv.asd.isc.noout[-1], 2, median)
isc.factor.snv = isc.subtracted.snv.median['1415'] # Water band 1400 nm or *1415nm

source.snv.factor.cor = source.snv

source.snv.factor.cor$NIR = source.snv.factor.cor$NIR + isc.factor.snv
```

#### All in one Calibration
```{R}
jpeg(paste0(dir.p, '/results/figs/factor_correction_nir_asd_isc_SNV_osc_calibration.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(wavelength, 
        colMeans(target.snv$NIR),
        #xlim = c(400, 2450),
        ylim = c(-3,3),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 3,
        lwd = 2,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.8),
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'SG-SNV',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(wavelength, 
        colMeans(source.snv$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)

matplot(wavelength, 
        colMeans(source.snv.factor.cor$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 1.0, green = 0.0, blue = 0.0, alpha = 0.5),
        las=1, font=2,
        add = T)

legend("topleft",bty="n", legend=c("Target: ASD data","Source: ISC NIRScan data", "Index: Fitting source in target data"),
       col=c(rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
             "forestgreen",
             "red"),
       lty=c(1),cex=1,lwd=3)

dev.off()

## VALIDATION
source.snv.factor.cor.val = source.snv.val

source.snv.factor.cor.val$NIR = source.snv.factor.cor.val$NIR + isc.factor.snv

## All in one Validation
jpeg(paste0(dir.p, '/results/figs/factor_correction_nir_asd_isc_SNV_osc_validation.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(wavelength, 
        colMeans(target.snv.val$NIR),
        #xlim = c(400, 2450),
        ylim = c(-3,3),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 3,
        lwd = 2,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.8),
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'Reflectance Factor (x100)',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(wavelength, 
        colMeans(source.snv.val$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)

matplot(wavelength, 
        colMeans(source.snv.factor.cor.val$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 1.0, green = 0.0, blue = 0.0, alpha = 0.5),
        las=1, font=2,
        add = T)

legend("topleft",bty="n", legend=c("Target: ASD data","Source: ISC NIRScan data", "Index: Fitting source in target data"),
       col=c(rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
             "forestgreen",
             "red"),
       lty=c(1),cex=1,lwd=3)
 
dev.off()
```


#### Exporting data.frame

```{R}
source.snv.factor.cor
write.csv(source.snv.factor.cor, paste0(dir.p, '/tables/db_model/db_source_spec_snv_subcorr_cal.csv'))
source.snv.factor.cor.val
write.csv(source.snv.factor.cor.val, paste0(dir.p, '/tables/db_model/db_source_spec_snv_subcorr_val.csv'))
```

### Piecewise Direct Standardisation (PDS) ----

#### Define the Piecewise Direct Standardization (PDS) function with PLS regression

```{R}
perform_pds <- function(target, source, MWsize, Ncomp, wavelength){
  
  require(pls)
  
  #Loop Initialization:
  i<-MWsize
  k<-i-1
  #Creation of an empty P matrix:
  P<-matrix(0,nrow=ncol(target),ncol=ncol(target)-(2*i)+2)
  InterceptReg<-c()
  
  while(i<=(ncol(target)-k)){
    
    #PLS regression:
    fit<- plsr(target[,i] ~ as.matrix(source[,(i-k):(i+k)]),
               ncomp=Ncomp, scale=F, method="oscorespls")
    
    #Extraction of the regression coefficients:
    coefReg<-as.numeric(coef(fit, ncomp=Ncomp, intercept = TRUE))
    InterceptReg<-c(InterceptReg,coefReg[1])
    coefReg<-coefReg[2:length(coefReg)]
    
    #Add coefficients to the transfer matrix:
    P[(i-k):(i+k),i-k]<-t(coefReg)
    
    rm(coefReg,fit)
    i<-i+1
    
    #Diplay progression:
    cat("\r",paste(round(i/ncol(target)*100)," %",sep=""))}
  
  P<-data.frame(matrix(0,nrow=ncol(target),ncol=k), P,
                matrix(0,nrow=ncol(target),ncol=k))
  InterceptReg<-c(rep(0,k),InterceptReg,rep(0,k)) 
  
  Output<-list(P = P , Intercept = InterceptReg)
  
  return(Output)}

# Apply PDS to transfer spectra from source to target instrument
Ncomp <- 2 # 3
MWsize <- 2 # 8
#wavelength = as.numeric(substring(colnames(target.snv$NIR), 1, 7))
transformed.pds.target <- perform_pds(target.snv$NIR, source.snv$NIR, MWsize, Ncomp, wavelength)

## Matrix that will help to transfer spec from source to target using SNV data
source.snv.pds.cal <- as.matrix(source.snv$NIR)%*%as.matrix(transformed.pds.target$P)
source.snv.pds.cal <- sweep(source.snv.pds.cal, 2, as.numeric(t(transformed.pds.target$Intercept)), "+")
source.snv.pds.cal[1:5, 1:141]
colnames(source.snv.pds.cal[, 1:141])
source.snv.pds.cal <- data.frame(source.snv$id, NIR = I(source.snv.pds.cal[, 2:140]))
colnames(source.snv.pds.cal)[1] = "id"
```

#### All in one graph - calibration (plotting)

```{R}
# jpeg(paste0(dir.p, '/results/figs/MEAN_SPEC_SIGN_pds_correction_nir_asd_isc_osc_cal.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(wavelength[2:140],
        colMeans(target.snv$NIR[,2:140]),
        #xlim = c(400, 2450),
        ylim = c(-3,3),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 3,
        lwd = 2,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.8), 
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'SG-SNV',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(wavelength[2:140],
        colMeans(source.snv$NIR[, 2:140]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.8),
        las=1, font=2,
        add = T)

matplot(wavelength[2:140],
        colMeans(source.snv.pds.cal$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 1.0, green = 0.0, blue = 0.0, alpha = 0.8),
        las=1, font=2,
        add = T)
legend("topleft",bty="n", legend=c("Target: ASD data","Source: ISC NIRScan data", "PDS: Fitting source in target data"),
       col=c(rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
             "forestgreen",
             "red"),
       lty=c(1),cex=1,lwd=3)

# dev.off()
```

### VALIDATION
Matrix that will help to transfer spec from source to target using SNV data

```{R}
source.snv.pds.val <- as.matrix(source.snv.val$NIR)%*%as.matrix(transformed.pds.target$P)
source.snv.pds.val <- sweep(source.snv.pds.val, 2, as.numeric(t(transformed.pds.target$Intercept)), "+")
source.snv.pds.val[1:5, 1:141]
colnames(source.snv.pds.val[, 1:141])
source.snv.pds.val <- data.frame(source.snv.val$id, NIR = I(source.snv.pds.val[, 2:140]))
colnames(source.snv.pds.val)[1] = "id"
```

#### All in one graph - calibration (plotting)
```{R}

# jpeg(paste0(dir.p, '/results/figs/MEAN_SPEC_SIGN_pds_correction_nir_asd_isc_osc_val.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(wavelength[2:140],
        colMeans(target.snv.val$NIR[,2:140]),
        #xlim = c(400, 2450),
        ylim = c(-3,3),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 3,
        lwd = 2,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.8), 
        xlab =  'Wavelength (nm)', #expression(paste('Wavenumber(cm'^'-1',')'))
        ylab = 'SG-SNV',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)

matplot(wavelength[2:140],
        colMeans(source.snv.val$NIR[, 2:140]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)

matplot(wavelength[2:140],
        colMeans(source.snv.pds.val$NIR),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 1.0, green = 0.0, blue = 0.0, alpha = 0.8),
        las=1, font=2,
        add = T)
legend("topleft",bty="n", legend=c("Target: ASD data","Source: ISC NIRScan data", "PDS: Fitting source in target data"),
       col=c(rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.3),
             "forestgreen",
             "red"),
       lty=c(1),cex=1,lwd=3)

# dev.off()
```

#### Exporting data.frame

```{r}
colnames(source.snv.pds.cal$NIR) = wavelength[2:140]
write.csv(source.snv.pds.cal, paste0(dir.p, '/tables/db_model/db_source_spec_snv_pdscorr_cal.csv'))

source.snv.pds.val

colnames(source.snv.pds.val$NIR) = wavelength[2:140]
write.csv(source.snv.pds.val, paste0(dir.p, '/tables/db_model/db_source_spec_snv_pdscorr_val.csv'))
```

### Other method ----

```{R}
```

### Saving R workspace ----

```{R}
save.image(file = paste0(dir.p, '/r_code/proj_low_cost_nir.RData'))
```
