---
title: "Data Analysis with R"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: tango
    fig_caption: true
    df_print: kable
    code_folding: hide
    code_download: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries

We load the necessary libraries. If they are not installed, we install them first.

```{r}
if(!require()){install.packages(''); require()}
if(!require(parallel)){install.packages('parallel'); require(parallel)}
if(!require(doParallel)){install.packages('doParallel'); require(doParallel)}
if(!require(pls)){install.packages('pls'); require(pls)}
if(!require(Cubist)){install.packages('Cubist'); require(Cubist)}
if(!require(caret)){install.packages('caret'); require(caret)}
if(!require(ggplot2)){install.packages('ggplot2'); require(ggplot2)}
```

### Good of Fitness (GOOF) to get model evaluation

This function calculates various model evaluation statistics such as Coefficient of determination, Standard error of prediction, Bias, Residual variance, Ratio of performance to deviation, Ratio of performance to interquartile distance, and Concordance.

This function, `goof`, calculates the goodness of fit between observed and predicted values. It computes various metrics such as the coefficient of determination (MEC), standard error of prediction (SEP), bias, residual variance (SEPc), ratio of performance to deviation (RPD), ratio of performance to interquartile distance (RPIQ), and concordance (ccc). If `plot.it` is set to `TRUE`, it also generates a scatter plot of observed vs predicted values. The `type` argument allows you to specify whether the output should be formatted for DSM or spectral data. The function returns a data frame with all the calculated metrics. Please note that this function requires the `eqscplot` function for equal-scaled plotting, which is not defined in this script. You might need to load a library (like `Hmisc`) or define this function before using `goof`.

```{r}
# Function to calculate goodness of fit
goof <- function(observed, predicted, plot.it = FALSE, type = "DSM") {
  
  # Coefficient of determination
  # This is calculated by fitting a linear model between predicted and observed values.
  rLM <- lm(predicted ~ observed)
  MEC <- as.matrix(summary(rLM)$adj.r.squared)
  
  # Standard error of prediction ^2
  # This is the mean squared difference between observed and predicted values.
  SEP2 <- mean((observed - predicted)^2)
  
  # Standard error of prediction
  # This is the square root of the standard error of prediction squared.
  SEP <- sqrt(SEP2)
  
  # Bias
  # This is the difference between the mean of predicted values and the mean of observed values.
  bias <- mean(predicted) - mean(observed)
  
  # Residual variance
  # This is the mean squared difference between observed values and predicted values corrected for bias.
  SEP2c <- sum(((predicted - bias - observed)^2) / length(observed))
  SEPc <- sqrt(SEP2c)
  
  # Ratio of performance to deviation
  # This is the ratio of the standard deviation of observed values to the standard error of prediction.
  RPD <- sd(observed) / SEP
  
  # Ratio of performance to interquartile distance
  # This is the ratio of the interquartile range of observed values to the standard error of prediction.
  IQ <- c(quantile(observed))[4] - c(quantile(observed))[2]
  RPIQ <- IQ / SEP
  
  # Concordance
  # This is a measure of the agreement between observed and predicted values.
  mx <- mean(observed)
  my <- mean(predicted)
  s2x <- var(observed)
  s2y <- var(predicted)
  sxy <- mean((observed - mx) * (predicted - my))
  ccc <- 2 * sxy / (s2x + s2y + (mx - my)^2)
  
  # If plot.it is TRUE, generate a scatter plot of observed vs predicted values.
  if (plot.it == TRUE) {
    eqscplot(observed, predicted)
    abline(a = 0, b = 1, col = "brown4")
  }
  
  # Depending on the type of output required, return a data frame with the calculated metrics.
  if (type == "DSM") { 
    gf <- data.frame(MEC = MEC, 
                     concordance = ccc, 
                     MSE = SEP2, 
                     RMSE = SEP, 
                     bias = bias, 
                     row.names = NULL)
  } else if (type == "spec") { 
    gf <- data.frame(MEC = MEC, 
                     concordance = ccc, 
                     MSE = SEP2, 
                     RMSE = SEP, 
                     bias = bias, 
                     MSEc = SEP2c, 
                     RMSEc = SEPc, 
                     RPD = RPD, 
                     RPIQ = RPIQ, 
                     row.names = NULL)
  } else { 
    stop("ERROR: Revise the type of output you require. Select from either DSM or spec")
  } 
  return(gf)
}
```

#### Setting the folder path ----

```{r}
dir.p = 'D:/project_lowcost_nir/database' # homeoffice/
```

#### Importing soil data ----

```{r}
db.soil = read.csv(paste0(dir.p, '/tables/db_model/db_soil_oscar.csv'))
db.soil
db.soil$id = sprintf("osc_%02d", db.soil$id)
db.soil = db.soil[1:102,]
db.soil = db.soil[!(db.soil$tn.perc == "#N/A" & db.soil$tc.perc == "#N/A"), ]
```

### 1st model: ASD vis-NIR range (400 - 2450 nm) ----

The **1st model: ASD vis-NIR range (400 - 2450 nm)** refers to a model that uses the visible and near-infrared (vis-NIR) spectral range of 400 to 2450 nanometers. This range is commonly used in remote sensing and spectroscopy applications, including soil analysis.

In the context of soil analysis, this model would use the reflectance values of soil samples across the specified vis-NIR range. These reflectance values are often obtained using a spectrometer, such as an Analytical Spectral Devices (ASD) spectrometer. The data is then processed and analyzed, often using statistical or machine learning techniques, to predict various soil properties.

This model is referred to as the "1st model" indicating that it's the initial or primary model used in the analysis process. It's important to note that the effectiveness and accuracy of this model can depend on various factors, including the quality of the spectral data, the preprocessing and analysis techniques used, and the specific soil properties being predicted.

Please note that the specific details and implementation of this model can vary based on the context and the specific requirements of the analysis. For example, different soil properties might be predicted using different spectral ranges or analysis techniques. Therefore, it's always important to understand the specific context and requirements when working with such models.

```{r}
db.asd.visnir.cal = read.csv(paste0(dir.p, '/tables/db_model/asd_visnir_sg_cal.csv'))[-1]
db.asd.visnir.cal[1:5, c(1:4, ncol(db.asd.visnir.cal))]
colnames(db.asd.visnir.cal) = gsub("X", "", colnames(db.asd.visnir.cal))
```

#### Filtering from 400 - 2450 nm

This script is performing a Partial Least Squares Regression (PLSR) analysis on soil spectral data in the visible and near-infrared (vis-NIR) range (400 - 2450 nm). The script starts by filtering the spectral data to include only the wavelengths between 400 nm and 2450 nm. It then merges this spectral data with the soil data. The target variables (total nitrogen and total carbon percentages) are converted to numeric format. The script then sets up parallel processing for the PLSR model, using all but two of the available cores.

```{r}
# Identify the column numbers for the wavelengths 400 nm and 2450 nm
which(colnames(db.asd.visnir.cal) == "400") 
which(colnames(db.asd.visnir.cal) == "2450") 

# Filter the data to include only the columns between 400 nm and 2450 nm
db.asd.visnir.cal = db.asd.visnir.cal[, c(1, 47:2097)]

# Merge the soil data with the spectral data
db.soil.asd.visnir.cal = merge(db.soil, db.asd.visnir.cal, by="id")
db.soil.asd.visnir.cal[1:5, c(1:10, ncol(db.soil.asd.visnir.cal))]

# Convert the target variables to numeric
db.soil.asd.visnir.cal$tn.perc = as.numeric(db.soil.asd.visnir.cal$tn.perc)
db.soil.asd.visnir.cal$tc.perc = as.numeric(db.soil.asd.visnir.cal$tc.perc)

# Detect the number of cores to use for parallel processing
cores = parallel::detectCores()-2
```

#### Partial Least Squares Regression (PLSR)

Partial Least Squares Regression (PLSR) is a statistical method that finds a linear regression model by projecting the predicted variables and the observable variables to a new space¹. It is particularly suited when the matrix of predictors has more variables than observations, and when there is multicollinearity among predictor values¹.

PLSR constructs new predictor variables, known as components, as linear combinations of the original predictor variables². Unlike Principal Components Regression (PCR) which creates components to explain the observed variability in the predictor variables without considering the response variable, PLSR does take the response variable into account. Therefore, it often leads to models that are able to fit the response variable with fewer components².

PLSR is widely used in various fields including chemometrics, bioinformatics, sensometrics, neuroscience, and anthropology¹. It is especially useful for modeling spectral data, such as in soil analysis³. The effectiveness and accuracy of PLSR can depend on various factors, including the quality of the spectral data, the preprocessing and analysis techniques used, and the specific properties being predicted¹².

Source: Conversation with Bing, 05/11/2023

(1) Partial least squares regression - Wikipedia. <https://en.wikipedia.org/wiki/Partial_least_squares_regression>.

(2) Partial Least Squares Regression and Principal Components ... - MathWorks. <https://www.mathworks.com/help/stats/partial-least-squares-regression-and-principal-components-regression.html>.

(3) Partial Least Squares Regression (PLSR) Applied to NIR and ... - Springer. <https://link.springer.com/article/10.1007/s12393-016-9147-1>.

(4) Partial Least-Squares Regression (PLSR) \| SpringerLink. <https://link.springer.com/referenceworkentry/10.1007/978-1-4419-9863-7_1274>.

(5) plsR function - RDocumentation. <https://www.rdocumentation.org/packages/plsRglm/versions/1.5.1/topics/plsR>.

The PLSR model is trained using cross-validation, and the time taken to train the model is measured. The script then stops the parallel processing and prints a summary of the PLSR model. Finally, the script selects the optimal number of components for the PLSR model based on the Prediction Error Sum of Squares (PRESS) statistic.

```{r}
# Select the columns to be used in the PLSR model
db.tc.asd.visnir.pls = db.soil.asd.visnir.cal[, c(6, 9:ncol(db.soil.asd.visnir.cal))]

# Set up parallel processing for the PLSR model
pls::pls.options(parallel = makeCluster(cores, type = 'PSOCK'))

# Train the PLSR model with cross-validation and measure the time taken
system.time(tc.asd.visnir.pls.m <- pls::plsr(tc.perc~., 
                                             ncomp = 15, 
                                             data = db.tc.asd.visnir.pls, 
                                             valid = 'CV', 
                                             scale = F)) 

# Stop the parallel processing
stopCluster(pls.options()$parallel)

# Print a summary of the PLSR model
summary(tc.asd.visnir.pls.m)

# Extract the variable importance from the PLSR model
tc.asd.visnir.pls.m$validation

# Select the optimal number of components based on the PRESS statistic
ncomp.sel = as.data.frame(tc.asd.visnir.pls.m$validation$PRESS)
colnames(ncomp.sel) = gsub("comps", "", names(ncomp.sel))
ncomp = as.numeric(colnames(ncomp.sel[which.min(ncomp.sel[1,])]))
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/ncomp/soc_asd_visnir_PLSR_Optimal_N_Comp.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
plot(RMSEP(tc.asd.visnir.pls.m),
     main = '',
     xlim = c(0, 15),
     xaxt = 'n',
     frame.plot = FALSE,
     type = 'l',
     lty = 1:2,
     col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
     xlab =  'Number of components', # expression(paste('Wavelength(cm'^'-1',')')),
     ylab = 'RMSEP',
     las=1, font=2)
axis(1, at = seq(0, 15, by = 1), labels = T)
mtext('SOC SG-0D')
legend("topright",legend=c("Cross-Validation (CV)","Adjusted CV"),
       lty = 1:2,
       col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
       bty = 'n',
       x.intersp = 0.25,
       seg.len = 1)
points(ncomp, 0.2076, col = 'red', pch =20)
# dev.off()
```

#### Predicting

This script is part of a larger process of analyzing soil properties using spectral data. The validation data is used to evaluate the performance of a predictive model that has been trained on a separate set of calibration data. The model uses the spectral data in the vis-NIR range to predict various soil properties. The effectiveness and accuracy of the model can be evaluated by comparing the model's predictions on the validation data to the actual observed values.

```{r}
# Read the validation data from a CSV file
db.asd.visnir.val = read.csv(paste0(dir.p, '/tables/db_model/asd_visnir_sg_val.csv'))[-1]

# Display the first 5 rows and the last column of the validation data
db.asd.visnir.val[1:5, c(1:4, ncol(db.asd.visnir.val))]

# Remove the "X" from the column names of the validation data
colnames(db.asd.visnir.val) = gsub("X", "", colnames(db.asd.visnir.val))
```

#### Filtering from 400 - 2450 nm (ASD)

```{r}
# Identify the column numbers for the wavelengths 400 nm and 2450 nm
which(colnames(db.asd.visnir.val) == "400") 
which(colnames(db.asd.visnir.val) == "2450") 

# Filter the data to include only the columns between 400 nm and 2450 nm
db.asd.visnir.val = db.asd.visnir.val[, c(1, 47:2097)]

# Merge the soil data with the spectral data
db.soil.asd.visnir.val = merge(db.soil, db.asd.visnir.val, by="id")

# Display the first 5 rows and the last 10 columns of the merged data
db.soil.asd.visnir.val[1:5, c(1:10, ncol(db.soil.asd.visnir.val))]
```

#### Creating a validation table for all results

This script is part of a larger process of analyzing soil properties using spectral data. The validation data is used to evaluate the performance of a predictive model that has been trained on a separate set of calibration data. The model uses the spectral data in the vis-NIR range to predict various soil properties. The effectiveness and accuracy of the model can be evaluated by comparing the model's predictions on the validation data to the actual observed values. The `goof` function is used to calculate various goodness of fit metrics, such as the coefficient of determination (MEC), standard error of prediction (SEP), bias, residual variance (SEPc), ratio of performance to deviation (RPD), ratio of performance to interquartile distance (RPIQ), and concordance (ccc).

```{r}
# Create a new data frame with the first 6 columns of the validation data
db.soil.val.all = as.data.frame(db.soil.asd.visnir.val[, 1:6])

# Display the structure of the new data frame
str(db.soil.val.all)

#write.csv(db.soil.val.all[,c(1, 5:6)], paste0(dir.p, '/tables/db_soil_asd_nir_val.csv'))

# Convert the total nitrogen and total carbon percentages to numeric format
db.soil.val.all$tn.perc = as.numeric(db.soil.val.all$tn.perc)
db.soil.val.all$tc.perc = as.numeric(db.soil.val.all$tc.perc)

# Predict the total carbon percentage using the PLSR model and the validation data
db.soil.val.all$tc.asd.visnir.pls = stats::predict(tc.asd.visnir.pls.m, 
                                                   newdata = db.asd.visnir.val[, 2:ncol(db.asd.visnir.val)],
                                                   ncomp = ncomp)

# Convert the predicted total carbon percentages to numeric format
db.soil.val.all$tc.asd.visnir.pls = as.numeric(db.soil.val.all$tc.asd.visnir.pls[, 1, 1])

# Calculate the goodness of fit between the observed and predicted total carbon percentages
metric.tc.asd.visnir.pls = goof(observed = db.soil.val.all$tc.perc, 
                           predicted = db.soil.val.all$tc.asd.visnir.pls,
                           type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/soc_sg_asd_visnir_PLSR_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd.visnir.pls', 'tc.perc')])
plot(db.soil.val.all$tc.asd.visnir.pls, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd.visnir.pls ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.asd.visnir.pls$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.asd.visnir.pls$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.asd.visnir.pls$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.asd.visnir.pls$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.asd.visnir.pls$RPIQ, 2))))
# dev.off()
```

#### The Cubist model

The Cubist model is a type of rule-based model that is particularly effective for regression tasks. It builds a series of rule-based models and combines them to make predictions. In this script, the Cubist model is being used to predict the total carbon percentage in soil based on spectral data in the vis-NIR range. The effectiveness and accuracy of the model can be evaluated by comparing the model's predictions on the validation data to the actual observed values. The `varImp` function is used to calculate the importance of each variable in the model, which can provide insights into which wavelengths are most informative for predicting the total carbon percentage.

```{r}
# Set up the control parameters for the model training process
ctrl = caret::trainControl(method = 'repeatedcv', 
                           number = 10, 
                           repeats = 10, 
                           search = 'random')

# Check the number of cores available for parallel processing
cores # it has the number of cores from parallel::detectCores()-2

# Create a cluster using all available cores minus two
cl = parallel::makeCluster(cores) # Create a cluster using all cores

# Register the cluster for parallel processing
doParallel::registerDoParallel(cl)

# Train a Cubist model
#colnames(db.soil.asd.visnir.cal)[1:20]
system.time(tc.asd.visnir.cub.m <- caret::train(x=db.soil.asd.visnir.cal[ ,9:ncol(db.soil.asd.visnir.cal)], 
                                                y=db.soil.asd.visnir.cal$tc.perc,
                                                method = "cubist", 
                                                trControl = ctrl,
                                                importance = T))  
# Stop the parallel processing
stopCluster(cl)

# Print a summary of the Cubist model
print(tc.asd.visnir.cub.m)

# Calculate the variable importance of the Cubist model
tc.asd.visnir.cub.m.var = varImp(tc.asd.visnir.cub.m, scale = T)
```

```{r}
# jpeg(paste0(dir.p, '/results/figs_cub/varimp/soc_sg_asd_visnir_Var_Importance_Cub.jpeg'), width = 10, height = 10, units = 'in', res = 300)
plot(tc.asd.visnir.cub.m.var, top = 20, main = "Original data")
# dev.off()
```

#### Predicting

```{r}
# Merge the validation data with the soil data
db.soil.asd.visnir.val = merge(db.soil, db.asd.visnir.val, by="id")

# Display the first 5 rows and the last 10 columns of the merged data
db.soil.asd.visnir.val[1:5, c(1:10, ncol(db.soil.asd.visnir.val))]

db.soil.asd.visnir.val[1:5,1:10]

# Predict the total carbon percentage using the Cubist model and the validation data
db.soil.val.all$tc.asd.visnir.cub = stats::predict(tc.asd.visnir.cub.m, 
                                                   db.soil.asd.visnir.val[, 9:ncol(db.soil.asd.visnir.val)]) # Predicting using external data

# Calculate the goodness of fit between the observed and predicted total carbon percentages
metric.tc.asd.visnir.cub = goof(observed = db.soil.val.all$tc.perc, 
                                predicted = db.soil.val.all$tc.asd.visnir.cub,
                                type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_cub/soc_sg_asd_visnir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd.visnir.cub', 'tc.perc')])
plot(db.soil.val.all$tc.asd.visnir.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd.visnir.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.asd.visnir.cub$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.asd.visnir.cub$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.asd.visnir.cub$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.asd.visnir.cub$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.asd.visnir.cub$RPIQ, 2))))
# dev.off()
```

### 2nd model: ASD NIR range (955 - 1645 nm) ----

```{r}
db.asd.nir.cal = read.csv(paste0(dir.p, '/tables/db_model/db_target_asd_spec_cal.csv'))[-1]
db.asd.nir.cal[1:5, c(1:4, ncol(db.asd.nir.cal))]
colnames(db.asd.nir.cal) = gsub("NIR.", "", colnames(db.asd.nir.cal))
which(colnames(db.asd.nir.cal)=="1645")
db.asd.nir.cal = db.asd.nir.cal[, c(1, 3:141)]

db.soil.asd.nir.cal = merge(db.soil, db.asd.nir.cal, by="id")
db.soil.asd.nir.cal[1:5, c(1:10, ncol(db.soil.asd.nir.cal))]
str(db.soil.asd.nir.cal[, 1:10])
db.soil.asd.nir.cal$tn.perc = as.numeric(db.soil.asd.nir.cal$tn.perc)
db.soil.asd.nir.cal$tc.perc = as.numeric(db.soil.asd.nir.cal$tc.perc)
```

#### Partial Least Squares Regression (PLSR)

```{r}
## PLSR
colnames(db.soil.asd.nir.cal)[1:20]
db.tc.asd.nir.pls = db.soil.asd.nir.cal[, c(6, 9:ncol(db.soil.asd.nir.cal))]
colnames(db.tc.asd.nir.pls)[1:20]

pls::pls.options(parallel = makeCluster(cores, type = 'PSOCK'))
system.time(tc.asd.nir.pls.m <- pls::plsr(tc.perc~., 
                                             ncomp = 15, 
                                             data = db.tc.asd.nir.pls, 
                                             valid = 'CV', 
                                             scale = F)) 

stopCluster(pls.options()$parallel)

summary(tc.asd.nir.pls.m)
tc.asd.nir.pls.m$validation
ncomp.sel = as.data.frame(tc.asd.nir.pls.m$validation$PRESS)
colnames(ncomp.sel) = gsub("comps", "", names(ncomp.sel))
ncomp = as.numeric(colnames(ncomp.sel[which.min(ncomp.sel[1,])]))
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/ncomp/soc_asd_nir_PLSR_Optimal_N_Comp.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
plot(RMSEP(tc.asd.nir.pls.m),
     main = '',
     xlim = c(0, 15),
     xaxt = 'n',
     frame.plot = FALSE,
     type = 'l',
     lty = 1:2,
     col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
     xlab =  'Number of components', # expression(paste('Wavelength(cm'^'-1',')')),
     ylab = 'RMSEP',
     las=1, font=2)
axis(1, at = seq(0, 15, by = 1), labels = T)
mtext('SOC SG-0D: ASD NIR')
legend("topright",legend=c("Cross-Validation (CV)","Adjusted CV"),
       lty = 1:2,
       col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
       bty = 'n',
       x.intersp = 0.25,
       seg.len = 1)
points(ncomp, 0.1718, col = 'red', pch =20)
# dev.off()
```

#### Predicting

```{r}
db.asd.nir.val = read.csv(paste0(dir.p, '/tables/db_model/db_target_asd_spec_val.csv'))[-1]
db.asd.nir.val[1:5, c(1:4, ncol(db.asd.nir.val))]
colnames(db.asd.nir.val) = gsub("NIR.", "", colnames(db.asd.nir.val))
which(colnames(db.asd.nir.val)=="1645")
db.asd.nir.val = db.asd.nir.val[, c(1, 3:141)]

db.soil.val.all$tc.asd.nir.pls = stats::predict(tc.asd.nir.pls.m, 
                                                   newdata = db.asd.nir.val[, 2:ncol(db.asd.nir.val)],
                                                   ncomp = ncomp)

db.soil.val.all$tc.asd.nir.pls = as.numeric(db.soil.val.all$tc.asd.nir.pls[, 1, 1])

metric.tc.asd.nir.pls = goof(observed = db.soil.val.all$tc.perc, 
                                predicted = db.soil.val.all$tc.asd.nir.pls,
                                type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/soc_sg_asd_nir_PLSR_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd.nir.pls', 'tc.perc')])
plot(db.soil.val.all$tc.asd.nir.pls, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd.nir.pls ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.asd.nir.pls$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.asd.nir.pls$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.asd.nir.pls$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.asd.nir.pls$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.asd.nir.pls$RPIQ, 2))))
# dev.off()
```

#### Cubist

```{r}
## Cubist
cores # it has the number of cores from parallel::detectCores()-2
cl = parallel::makeCluster(cores) # Create a cluster using all cores
doParallel::registerDoParallel(cl)

system.time(tc.asd.nir.cub.m <- caret::train(x=db.soil.asd.nir.cal[ ,9:ncol(db.soil.asd.nir.cal)], 
                                                y=db.soil.asd.nir.cal$tc.perc,
                                                method = "cubist", 
                                                trControl = ctrl,
                                                importance = T))  
stopCluster(cl)

print(tc.asd.nir.cub.m)

tc.asd.nir.cub.m.var = varImp(tc.asd.nir.cub.m, scale = T)
```

```{r}
# jpeg(paste0(dir.p, '/results/figs_cub/varimp/soc_sg_asd_nir_Var_Importance_Cub.jpeg'), width = 10, height = 10, units = 'in', res = 300)
plot(tc.asd.nir.cub.m.var, top = 20, main = "Original data")
# dev.off()
```

#### Predicting

```{r}

## Predicting
db.soil.val.all$tc.asd.nir.cub = stats::predict(tc.asd.nir.cub.m, 
                                                   db.asd.nir.val[, 2:ncol(db.asd.nir.val)]) # Predicting using external data

metric.tc.asd.nir.cub = goof(observed = db.soil.val.all$tc.perc, 
                                predicted = db.soil.val.all$tc.asd.nir.cub,
                                type = 'spec')


jpeg(paste0(dir.p, '/Results/figs_cub/soc_sg_asd_nir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd.nir.cub', 'tc.perc')])
plot(db.soil.val.all$tc.asd.nir.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd.nir.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.asd.nir.cub$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.asd.nir.cub$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.asd.nir.cub$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.asd.nir.cub$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.asd.nir.cub$RPIQ, 2))))
dev.off()
```

# 3rd model: ISC NIRScan NIR range (955 - 1645 nm) ----

```{r}
db.isc.nir.cal = read.csv(paste0(dir.p, '/tables/db_model/db_source_isc_spec_cal.csv'))[-1]
db.isc.nir.cal[1:5, c(1:4, ncol(db.isc.nir.cal))]
colnames(db.isc.nir.cal) = gsub("NIR.", "", colnames(db.isc.nir.cal))
which(colnames(db.isc.nir.cal) == "1645")
db.isc.nir.cal = db.isc.nir.cal[, c(1, 3:141)]

db.soil.isc.nir.cal = merge(db.soil, db.isc.nir.cal, by="id")
db.soil.isc.nir.cal[1:5, c(1:10, ncol(db.soil.isc.nir.cal))]
str(db.soil.isc.nir.cal[, 1:10])
db.soil.isc.nir.cal$tn.perc = as.numeric(db.soil.isc.nir.cal$tn.perc)
db.soil.isc.nir.cal$tc.perc = as.numeric(db.soil.isc.nir.cal$tc.perc)
```

```{r}
## PLSR
colnames(db.soil.isc.nir.cal)[1:20]
db.tc.isc.nir.pls = db.soil.isc.nir.cal[, c(6, 9:ncol(db.soil.isc.nir.cal))]
colnames(db.tc.isc.nir.pls)[1:20]

pls::pls.options(parallel = makeCluster(cores, type = 'PSOCK'))
system.time(tc.isc.nir.pls.m <- pls::plsr(tc.perc~., 
                                          ncomp = 15, 
                                          data = db.tc.isc.nir.pls, 
                                          valid = 'CV', 
                                          scale = F)) 

stopCluster(pls.options()$parallel)

summary(tc.isc.nir.pls.m)
tc.isc.nir.pls.m$validation
ncomp.sel = as.data.frame(tc.isc.nir.pls.m$validation$PRESS)
colnames(ncomp.sel) = gsub("comps", "", names(ncomp.sel))
ncomp = as.numeric(colnames(ncomp.sel[which.min(ncomp.sel[1,])]))
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/ncomp/soc_isc_nir_PLSR_Optimal_N_Comp.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
plot(RMSEP(tc.isc.nir.pls.m),
     main = '',
     xlim = c(0, 15),
     xaxt = 'n',
     frame.plot = FALSE,
     type = 'l',
     lty = 1:2,
     col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
     xlab =  'Number of components', # expression(paste('Wavelength(cm'^'-1',')')),
     ylab = 'RMSEP',
     las=1, font=2)
axis(1, at = seq(0, 15, by = 1), labels = T)
mtext('SOC SG-0D: isc NIR')
legend("topright",legend=c("Cross-Validation (CV)","Adjusted CV"),
       lty = 1:2,
       col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
       bty = 'n',
       x.intersp = 0.25,
       seg.len = 1)
points(ncomp, 0.3211, col = 'red', pch =20)
# dev.off()
```

```{r}
## Predicting
db.isc.nir.val = read.csv(paste0(dir.p, '/tables/db_model/db_source_isc_spec_val.csv'))[-1]
db.isc.nir.val[1:5, c(1:4, ncol(db.isc.nir.val))]
colnames(db.isc.nir.val) = gsub("NIR.", "", colnames(db.isc.nir.val))
which(colnames(db.isc.nir.val)== "1645")
db.isc.nir.val = db.isc.nir.val[, c(1, 3:141)]

db.soil.val.all$tc.isc.nir.pls = stats::predict(tc.isc.nir.pls.m, 
                                                newdata = db.isc.nir.val[, 2:ncol(db.isc.nir.val)],
                                                ncomp = ncomp)

db.soil.val.all$tc.isc.nir.pls = as.numeric(db.soil.val.all$tc.isc.nir.pls[, 1, 1])

metric.tc.isc.nir.pls = goof(observed = db.soil.val.all$tc.perc, 
                             predicted = db.soil.val.all$tc.isc.nir.pls,
                             type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/soc_sg_isc_nir_PLSR_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.nir.pls', 'tc.perc')])
plot(db.soil.val.all$tc.isc.nir.pls, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.nir.pls ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.isc.nir.pls$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.isc.nir.pls$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.isc.nir.pls$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.isc.nir.pls$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.isc.nir.pls$RPIQ, 2))))
# dev.off()
```

```{r}
## Cubist
cores # it has the number of cores from parallel::detectCores()-2
cl = parallel::makeCluster(cores) # Create a cluster using all cores
doParallel::registerDoParallel(cl)

system.time(tc.isc.nir.cub.m <- caret::train(x=db.soil.isc.nir.cal[ ,9:ncol(db.soil.isc.nir.cal)], 
                                             y=db.soil.isc.nir.cal$tc.perc,
                                             method = "cubist", 
                                             trControl = ctrl,
                                             importance = T))  
stopCluster(cl)

print(tc.isc.nir.cub.m)

tc.isc.nir.cub.m.var = varImp(tc.isc.nir.cub.m, scale = T)

# jpeg(paste0(dir.p, '/results/figs_cub/varimp/soc_sg_isc_nir_Var_Importance_Cub.jpeg'), width = 10, height = 10, units = 'in', res = 300)
plot(tc.isc.nir.cub.m.var, top = 20, main = "Original data")
# dev.off()
```

```{r}

## Predicting
db.soil.val.all$tc.isc.nir.cub = stats::predict(tc.isc.nir.cub.m, 
                                                db.isc.nir.val[, 2:ncol(db.isc.nir.val)]) # Predicting using external data

metric.tc.isc.nir.cub = goof(observed = db.soil.val.all$tc.perc, 
                             predicted = db.soil.val.all$tc.isc.nir.cub,
                             type = 'spec')


# jpeg(paste0(dir.p, '/Results/figs_cub/soc_sg_isc_nir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.nir.cub', 'tc.perc')])
plot(db.soil.val.all$tc.isc.nir.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.nir.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.isc.nir.cub$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.isc.nir.cub$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.isc.nir.cub$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.isc.nir.cub$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.isc.nir.cub$RPIQ, 2))))
# dev.off()
```

# 4th model: ASD SNV NIR range (955 - 1645 nm) ----

```{r}
db.asd.snv.nir.cal = read.csv(paste0(dir.p, '/tables/db_model/db_target_asd_spec_snv_cal.csv'))[-1]
db.asd.snv.nir.cal[1:5, c(1:4, ncol(db.asd.snv.nir.cal))]
colnames(db.asd.snv.nir.cal) = gsub("NIR.", "", colnames(db.asd.snv.nir.cal))
which(colnames(db.asd.snv.nir.cal)=="1645")
db.asd.snv.nir.cal = db.asd.snv.nir.cal[, c(1, 3:141)]

db.soil.asd.snv.nir.cal = merge(db.soil, db.asd.snv.nir.cal, by="id")
db.soil.asd.snv.nir.cal[1:5, c(1:10, ncol(db.soil.asd.snv.nir.cal))]
str(db.soil.asd.snv.nir.cal[, 1:10])
db.soil.asd.snv.nir.cal$tn.perc = as.numeric(db.soil.asd.snv.nir.cal$tn.perc)
db.soil.asd.snv.nir.cal$tc.perc = as.numeric(db.soil.asd.snv.nir.cal$tc.perc)
```

```{r}
## PLSR
colnames(db.soil.asd.snv.nir.cal)[1:20]
db.tc.asd.snv.nir.pls = db.soil.asd.snv.nir.cal[, c(6, 9:ncol(db.soil.asd.snv.nir.cal))]
colnames(db.tc.asd.snv.nir.pls)[1:20]

pls::pls.options(parallel = makeCluster(cores, type = 'PSOCK'))
system.time(tc.asd.snv.nir.pls.m <- pls::plsr(tc.perc~., 
                                          ncomp = 15, 
                                          data = db.tc.asd.snv.nir.pls, 
                                          valid = 'CV', 
                                          scale = F)) 

stopCluster(pls.options()$parallel)

summary(tc.asd.snv.nir.pls.m)
tc.asd.snv.nir.pls.m$validation
ncomp.sel = as.data.frame(tc.asd.snv.nir.pls.m$validation$PRESS)
colnames(ncomp.sel) = gsub("comps", "", names(ncomp.sel))
ncomp = as.numeric(colnames(ncomp.sel[which.min(ncomp.sel[1,])]))
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/ncomp/soc_asd_SNV_nir_PLSR_Optimal_N_Comp.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
plot(RMSEP(tc.asd.snv.nir.pls.m),
     main = '',
     xlim = c(0, 15),
     xaxt = 'n',
     frame.plot = FALSE,
     type = 'l',
     lty = 1:2,
     col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
     xlab =  'Number of components', # expression(paste('Wavelength(cm'^'-1',')')),
     ylab = 'RMSEP',
     las=1, font=2)
axis(1, at = seq(0, 15, by = 1), labels = T)
mtext('SOC SG-0D: ASD NIR')
legend("topright",legend=c("Cross-Validation (CV)","Adjusted CV"),
       lty = 1:2,
       col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
       bty = 'n',
       x.intersp = 0.25,
       seg.len = 1)
points(ncomp, 0.1616, col = 'red', pch =20)
# dev.off()
```

```{r}
## Predicting
db.asd.snv.nir.val = read.csv(paste0(dir.p, '/tables/db_model/db_target_asd_spec_snv_val.csv'))[-1]
db.asd.snv.nir.val[1:5, c(1:4, ncol(db.asd.snv.nir.val))]
colnames(db.asd.snv.nir.val) = gsub("NIR.", "", colnames(db.asd.snv.nir.val))
which(colnames(db.asd.snv.nir.val)=="1645")
db.asd.snv.nir.val = db.asd.snv.nir.val[, c(1, 3:141)]

db.soil.val.all$tc.asd.snv.nir.pls = stats::predict(tc.asd.snv.nir.pls.m, 
                                                newdata = db.asd.snv.nir.val[, 2:ncol(db.asd.snv.nir.val)],
                                                ncomp = ncomp)

db.soil.val.all$tc.asd.snv.nir.pls = as.numeric(db.soil.val.all$tc.asd.snv.nir.pls[, 1, 1])

metric.tc.asd.snv.nir.pls = goof(observed = db.soil.val.all$tc.perc, 
                             predicted = db.soil.val.all$tc.asd.snv.nir.pls,
                             type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/soc_SNV_asd_nir_PLSR_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd.snv.nir.pls', 'tc.perc')])
plot(db.soil.val.all$tc.asd.snv.nir.pls, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd.snv.nir.pls ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.asd.snv.nir.pls$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.asd.snv.nir.pls$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.asd.snv.nir.pls$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.asd.snv.nir.pls$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.asd.snv.nir.pls$RPIQ, 2))))
# dev.off()
```

```{r}
## Cubist
cores # it has the number of cores from parallel::detectCores()-2
cl = parallel::makeCluster(cores) # Create a cluster using all cores
doParallel::registerDoParallel(cl)

system.time(tc.asd.snv.nir.cub.m <- caret::train(x=db.soil.asd.snv.nir.cal[ ,9:ncol(db.soil.asd.snv.nir.cal)], 
                                             y=db.soil.asd.snv.nir.cal$tc.perc,
                                             method = "cubist", 
                                             trControl = ctrl,
                                             importance = T))  
stopCluster(cl)

print(tc.asd.snv.nir.cub.m)

tc.asd.snv.nir.cub.m.var = varImp(tc.asd.snv.nir.cub.m, scale = T)

# jpeg(paste0(dir.p, '/results/figs_cub/varimp/soc_SNV_asd_nir_Var_Importance_Cub.jpeg'), width = 10, height = 10, units = 'in', res = 300)
plot(tc.asd.snv.nir.cub.m.var, top = 20, main = "Original data")
# dev.off()
```

```{r}

## Predicting
db.soil.val.all$tc.asd.snv.nir.cub = stats::predict(tc.asd.snv.nir.cub.m, 
                                                db.asd.snv.nir.val[, 2:ncol(db.asd.snv.nir.val)]) # Predicting using external data

metric.tc.asd.snv.nir.cub = goof(observed = db.soil.val.all$tc.perc, 
                             predicted = db.soil.val.all$tc.asd.snv.nir.cub,
                             type = 'spec')

```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_cub/soc_SNV_asd_nir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd.snv.nir.cub', 'tc.perc')])
plot(db.soil.val.all$tc.asd.snv.nir.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd.snv.nir.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.asd.snv.nir.cub$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.asd.snv.nir.cub$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.asd.snv.nir.cub$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.asd.snv.nir.cub$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.asd.snv.nir.cub$RPIQ, 2))))
# dev.off()
```

# 5th model: ISC NIRScan SNV NIR range (955 - 1645 nm) ----

```{r}
db.isc.snv.nir.cal = read.csv(paste0(dir.p, '/tables/db_model/db_source_isc_spec_snv_cal.csv'))[-1]
db.isc.snv.nir.cal[1:5, c(1:4, ncol(db.isc.snv.nir.cal))]
colnames(db.isc.snv.nir.cal) = gsub("NIR.", "", colnames(db.isc.snv.nir.cal))
which(colnames(db.isc.snv.nir.cal)=="1645")
db.isc.snv.nir.cal = db.isc.snv.nir.cal[, c(1, 3:141)]

db.soil.isc.snv.nir.cal = merge(db.soil, db.isc.snv.nir.cal, by="id")
db.soil.isc.snv.nir.cal[1:5, c(1:10, ncol(db.soil.isc.snv.nir.cal))]
str(db.soil.isc.snv.nir.cal[, 1:10])
db.soil.isc.snv.nir.cal$tn.perc = as.numeric(db.soil.isc.snv.nir.cal$tn.perc)
db.soil.isc.snv.nir.cal$tc.perc = as.numeric(db.soil.isc.snv.nir.cal$tc.perc)
```

```{r}
## PLSR
colnames(db.soil.isc.snv.nir.cal)[1:20]
db.tc.isc.snv.nir.pls = db.soil.isc.snv.nir.cal[, c(6, 9:ncol(db.soil.isc.snv.nir.cal))]
colnames(db.tc.isc.snv.nir.pls)[1:20]

pls::pls.options(parallel = makeCluster(cores, type = 'PSOCK'))
system.time(tc.isc.snv.nir.pls.m <- pls::plsr(tc.perc~., 
                                              ncomp = 15, 
                                              data = db.tc.isc.snv.nir.pls, 
                                              valid = 'CV', 
                                              scale = F)) 

stopCluster(pls.options()$parallel)

summary(tc.isc.snv.nir.pls.m)
tc.isc.snv.nir.pls.m$validation
ncomp.sel = as.data.frame(tc.isc.snv.nir.pls.m$validation$PRESS)
colnames(ncomp.sel) = gsub("comps", "", names(ncomp.sel))
ncomp = as.numeric(colnames(ncomp.sel[which.min(ncomp.sel[1,])]))
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/ncomp/soc_isc_SNV_nir_PLSR_Optimal_N_Comp.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
plot(RMSEP(tc.isc.snv.nir.pls.m),
     main = '',
     xlim = c(0, 15),
     xaxt = 'n',
     frame.plot = FALSE,
     type = 'l',
     lty = 1:2,
     col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
     xlab =  'Number of components', # expression(paste('Wavelength(cm'^'-1',')')),
     ylab = 'RMSEP',
     las=1, font=2)
axis(1, at = seq(0, 15, by = 1), labels = T)
mtext('SOC SG-0D: isc NIR')
legend("topright",legend=c("Cross-Validation (CV)","Adjusted CV"),
       lty = 1:2,
       col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
       bty = 'n',
       x.intersp = 0.25,
       seg.len = 1)
points(ncomp, 0.1616, col = 'red', pch =20)
# dev.off()
```

```{r}
## Predicting
db.isc.snv.nir.val = read.csv(paste0(dir.p, '/tables/db_model/db_source_isc_spec_snv_val.csv'))[-1]
db.isc.snv.nir.val[1:5, c(1:4, ncol(db.isc.snv.nir.val))]
colnames(db.isc.snv.nir.val) = gsub("NIR.", "", colnames(db.isc.snv.nir.val))
which(colnames(db.isc.snv.nir.val)=="1645")
db.isc.snv.nir.val = db.isc.snv.nir.val[, c(1, 3:141)]

db.soil.val.all$tc.isc.snv.nir.pls = stats::predict(tc.isc.snv.nir.pls.m, 
                                                    newdata = db.isc.snv.nir.val[, 2:ncol(db.isc.snv.nir.val)],
                                                    ncomp = ncomp)

db.soil.val.all$tc.isc.snv.nir.pls = as.numeric(db.soil.val.all$tc.isc.snv.nir.pls[, 1, 1])

metric.tc.isc.snv.nir.pls = goof(observed = db.soil.val.all$tc.perc, 
                                 predicted = db.soil.val.all$tc.isc.snv.nir.pls,
                                 type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/soc_SNV_isc_nir_PLSR_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.snv.nir.pls', 'tc.perc')])
plot(db.soil.val.all$tc.isc.snv.nir.pls, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.snv.nir.pls ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.isc.snv.nir.pls$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.isc.snv.nir.pls$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.isc.snv.nir.pls$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.isc.snv.nir.pls$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.isc.snv.nir.pls$RPIQ, 2))))
# dev.off()
```

```{r}
## Cubist
cores # it has the number of cores from parallel::detectCores()-2
cl = parallel::makeCluster(cores) # Create a cluster using all cores
doParallel::registerDoParallel(cl)

system.time(tc.isc.snv.nir.cub.m <- caret::train(x=db.soil.isc.snv.nir.cal[ ,9:ncol(db.soil.isc.snv.nir.cal)], 
                                                 y=db.soil.isc.snv.nir.cal$tc.perc,
                                                 method = "cubist", 
                                                 trControl = ctrl,
                                                 importance = T))  
stopCluster(cl)

print(tc.isc.snv.nir.cub.m)

tc.isc.snv.nir.cub.m.var = varImp(tc.isc.snv.nir.cub.m, scale = T)
```

```{r}
# jpeg(paste0(dir.p, '/results/figs_cub/varimp/soc_SNV_isc_nir_Var_Importance_Cub.jpeg'), width = 10, height = 10, units = 'in', res = 300)
plot(tc.isc.snv.nir.cub.m.var, top = 20, main = "Original data")
# dev.off()
```

#### Predicting

```{r}
db.soil.val.all$tc.isc.snv.nir.cub = stats::predict(tc.isc.snv.nir.cub.m, 
                                                    db.isc.snv.nir.val[, 2:ncol(db.isc.snv.nir.val)]) # Predicting using external data

metric.tc.isc.snv.nir.cub = goof(observed = db.soil.val.all$tc.perc, 
                                 predicted = db.soil.val.all$tc.isc.snv.nir.cub,
                                 type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_cub/soc_SNV_isc_nir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.snv.nir.cub', 'tc.perc')])
plot(db.soil.val.all$tc.isc.snv.nir.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 2),
     ylim = c(0, 2),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.snv.nir.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(1.8, 0.5, bquote(RMSE == .(round(metric.tc.isc.snv.nir.cub$RMSE, 2))))
text(1.8, 0.4, bquote(MEC == .(round(metric.tc.isc.snv.nir.cub$MEC, 2))))
text(1.8, 0.3, bquote(CCC == .(round(metric.tc.isc.snv.nir.cub$concordance, 2))))
text(1.8, 0.2, bquote(bias == .(round(metric.tc.isc.snv.nir.cub$bias, 2))))
text(1.8, 0.1, bquote(RPIQ == .(round(metric.tc.isc.snv.nir.cub$RPIQ, 2))))
# dev.off()
```

### Predicting Transformed validation spectra within the best models ----

```{r}
## Substracted correction SG
db.isc.nir.subcorr.val = read.csv(paste0(dir.p, '/tables/db_model/db_source_spec_subcorr_val.csv'))[-1]
db.isc.nir.subcorr.val[1:5, c(1:4, ncol(db.isc.nir.subcorr.val))]
colnames(db.isc.nir.subcorr.val) = gsub("NIR.", "", colnames(db.isc.nir.subcorr.val))
which(colnames(db.isc.nir.subcorr.val)=="1645")
db.isc.nir.subcorr.val = db.isc.nir.subcorr.val[, c(1, 3:141)]
```

### 2nd model PLSR was the best fit with 15 pc.

```{r}
db.soil.val.all$tc.isc.nir.subcorr.pls = stats::predict(tc.asd.nir.pls.m,
                                                        newdata = db.isc.nir.subcorr.val[, 2:ncol(db.isc.nir.subcorr.val)],
                                                        ncomp = 15)

db.soil.val.all$tc.isc.nir.subcorr.pls = as.numeric(db.soil.val.all$tc.isc.nir.subcorr.pls[, 1, 1])

metric.tc.isc.nir.subcorr.pls = goof(observed = db.soil.val.all$tc.perc, 
                                 predicted = db.soil.val.all$tc.isc.nir.subcorr.pls,
                                 type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_pls/soc_isc_nir_SUBCORRECTED_PLSR_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.nir.subcorr.pls', 'tc.perc')])
plot(db.soil.val.all$tc.isc.nir.subcorr.pls, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 3),
     ylim = c(0, 3),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.nir.subcorr.pls ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(0.4, 2.9, bquote(RMSE == .(round(metric.tc.isc.nir.subcorr.pls$RMSE, 2))))
text(0.4, 2.8, bquote(MEC == .(round(metric.tc.isc.nir.subcorr.pls$MEC, 2))))
text(0.4, 2.7, bquote(CCC == .(round(metric.tc.isc.nir.subcorr.pls$concordance, 2))))
text(0.4, 2.6, bquote(bias == .(round(metric.tc.isc.nir.subcorr.pls$bias, 2))))
text(0.4, 2.5, bquote(RPIQ == .(round(metric.tc.isc.nir.subcorr.pls$RPIQ, 2))))
# dev.off()
```

### Substracted correction SG-SNV

```{r}
db.isc.nir.snv.subcorr.val = read.csv(paste0(dir.p, '/tables/db_model/db_source_spec_snv_subcorr_val.csv'))[-1]
db.isc.nir.snv.subcorr.val[1:5, c(1:4, ncol(db.isc.nir.snv.subcorr.val))]
colnames(db.isc.nir.snv.subcorr.val) = gsub("NIR.", "", colnames(db.isc.nir.snv.subcorr.val))
which(colnames(db.isc.nir.snv.subcorr.val)=="1645")
db.isc.nir.snv.subcorr.val = db.isc.nir.snv.subcorr.val[, c(1, 3:141)]
```

### 4th model Cubist was the best fit

```{r}
db.soil.val.all$tc.isc.nir.snv.subcorr.cub = stats::predict(tc.asd.snv.nir.cub.m,
                                                        newdata = db.isc.nir.snv.subcorr.val[, 2:ncol(db.isc.nir.snv.subcorr.val)])

metric.tc.isc.nir.snv.subcorr.cub = goof(observed = db.soil.val.all$tc.perc, 
                                     predicted = db.soil.val.all$tc.isc.nir.snv.subcorr.cub,
                                     type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_cub/soc_isc_nir_snv_SUBCORRECTED_CUB_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.nir.snv.subcorr.cub', 'tc.perc')])
plot(db.soil.val.all$tc.isc.nir.snv.subcorr.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 3),
     ylim = c(0, 3),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.nir.snv.subcorr.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(0.4, 2.9, bquote(RMSE == .(round(metric.tc.isc.nir.snv.subcorr.cub$RMSE, 2))))
text(0.4, 2.8, bquote(MEC == .(round(metric.tc.isc.nir.snv.subcorr.cub$MEC, 2))))
text(0.4, 2.7, bquote(CCC == .(round(metric.tc.isc.nir.snv.subcorr.cub$concordance, 2))))
text(0.4, 2.6, bquote(bias == .(round(metric.tc.isc.nir.snv.subcorr.cub$bias, 2))))
text(0.4, 2.5, bquote(RPIQ == .(round(metric.tc.isc.nir.snv.subcorr.cub$RPIQ, 2))))
# dev.off()
```

```{r}
## PDS correction SG-SNV
db.isc.nir.snv.pdscorr.val = read.csv(paste0(dir.p, '/tables/db_model/db_source_spec_snv_pdscorr_val.csv'))[-1]
db.isc.nir.snv.pdscorr.val[1:5, c(1:4, ncol(db.isc.nir.snv.pdscorr.val))]
colnames(db.isc.nir.snv.pdscorr.val) = gsub("NIR.", "", colnames(db.isc.nir.snv.pdscorr.val))
```

### 4th model Cubist was the best fit

```{r}
db.soil.val.all$tc.isc.nir.snv.pdscorr.cub = stats::predict(tc.asd.snv.nir.cub.m,
                                                            newdata = db.isc.nir.snv.pdscorr.val[, 2:ncol(db.isc.nir.snv.pdscorr.val)])

metric.tc.isc.nir.snv.pdscorr.cub = goof(observed = db.soil.val.all$tc.perc, 
                                         predicted = db.soil.val.all$tc.isc.nir.snv.pdscorr.cub,
                                         type = 'spec')
```

```{r}
# jpeg(paste0(dir.p, '/Results/figs_cub/soc_isc_nir_snv_PDSCORRECTED_CUB_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc.nir.snv.pdscorr.cub', 'tc.perc')])
plot(db.soil.val.all$tc.isc.nir.snv.pdscorr.cub, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 3),
     ylim = c(0, 3),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.isc.nir.snv.pdscorr.cub ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
text(0.4, 2.9, bquote(RMSE == .(round(metric.tc.isc.nir.snv.pdscorr.cub$RMSE, 2))))
text(0.4, 2.8, bquote(MEC == .(round(metric.tc.isc.nir.snv.pdscorr.cub$MEC, 2))))
text(0.4, 2.7, bquote(CCC == .(round(metric.tc.isc.nir.snv.pdscorr.cub$concordance, 2))))
text(0.4, 2.6, bquote(bias == .(round(metric.tc.isc.nir.snv.pdscorr.cub$bias, 2))))
text(0.4, 2.5, bquote(RPIQ == .(round(metric.tc.isc.nir.snv.pdscorr.cub$RPIQ, 2))))
# dev.off()
```

### Rounding table -----

```{r}
# Round the specified columns in the dataframe
colnames(db.soil.val.all)
db.soil.val.all[, 5:ncol(db.soil.val.all)] = round(db.soil.val.all[, 5:ncol(db.soil.val.all)], 2)
write.csv(db.soil.val.all, paste0(dir.p, '/tables/db_carbon_nitrogen_val.csv'))
```

### Saving R workspace ----

```{r}
save.image(file = paste0(dir.p, '/r_code/proj_low_cost_nir_modelling.RData'))
```
