---
title: "Rmarkdown for Soil Spectral Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Libraries
The following libraries are required for the analysis:

‘parallel’ and ‘doParallel’ for parallel computing,
‘pls’ and ‘Cubist’ for Partial Least Squares and Cubist regression,
‘caret’ for Classification And REgression Training,
‘ggplot2’ for creating elegant data visualisations.

```{r setup, include=FALSE}
if(!require()){install.packages(''); require()}
if(!require(parallel)){install.packages('parallel'); require(parallel)}
if(!require(doParallel)){install.packages('doParallel'); require(doParallel)}
if(!require(pls)){install.packages('pls'); require(pls)}
if(!require(Cubist)){install.packages('Cubist'); require(Cubist)}
if(!require(caret)){install.packages('caret'); require(caret)}
if(!require(ggplot2)){install.packages('ggplot2'); require(ggplot2)}
```

### Good of Fitness (GOOF) to get model evaluation
This function calculates various model evaluation statistics such as Coefficient of determination, Standard error of prediction, Bias, Residual variance, Ratio of performance to deviation, Ratio of performance to interquartile distance, and Concordance.

This function, `goof`, calculates the goodness of fit between observed and predicted values. It computes various metrics such as the coefficient of determination (MEC), standard error of prediction (SEP), bias, residual variance (SEPc), ratio of performance to deviation (RPD), ratio of performance to interquartile distance (RPIQ), and concordance (ccc). If `plot.it` is set to `TRUE`, it also generates a scatter plot of observed vs predicted values. The `type` argument allows you to specify whether the output should be formatted for DSM or spectral data. The function returns a data frame with all the calculated metrics. Please note that this function requires the `eqscplot` function for equal-scaled plotting, which is not defined in this script. You might need to load a library (like `Hmisc`) or define this function before using `goof`.

```{r setup, include=FALSE}
# Function to calculate goodness of fit
goof <- function(observed, predicted, plot.it = FALSE, type = "DSM") {
  
  # Coefficient of determination
  # This is calculated by fitting a linear model between predicted and observed values.
  rLM <- lm(predicted ~ observed)
  MEC <- as.matrix(summary(rLM)$adj.r.squared)
  
  # Standard error of prediction ^2
  # This is the mean squared difference between observed and predicted values.
  SEP2 <- mean((observed - predicted)^2)
  
  # Standard error of prediction
  # This is the square root of the standard error of prediction squared.
  SEP <- sqrt(SEP2)
  
  # Bias
  # This is the difference between the mean of predicted values and the mean of observed values.
  bias <- mean(predicted) - mean(observed)
  
  # Residual variance
  # This is the mean squared difference between observed values and predicted values corrected for bias.
  SEP2c <- sum(((predicted - bias - observed)^2) / length(observed))
  SEPc <- sqrt(SEP2c)
  
  # Ratio of performance to deviation
  # This is the ratio of the standard deviation of observed values to the standard error of prediction.
  RPD <- sd(observed) / SEP
  
  # Ratio of performance to interquartile distance
  # This is the ratio of the interquartile range of observed values to the standard error of prediction.
  IQ <- c(quantile(observed))[4] - c(quantile(observed))[2]
  RPIQ <- IQ / SEP
  
  # Concordance
  # This is a measure of the agreement between observed and predicted values.
  mx <- mean(observed)
  my <- mean(predicted)
  s2x <- var(observed)
  s2y <- var(predicted)
  sxy <- mean((observed - mx) * (predicted - my))
  ccc <- 2 * sxy / (s2x + s2y + (mx - my)^2)
  
  # If plot.it is TRUE, generate a scatter plot of observed vs predicted values.
  if (plot.it == TRUE) {
    eqscplot(observed, predicted)
    abline(a = 0, b = 1, col = "brown4")
  }
  
  # Depending on the type of output required, return a data frame with the calculated metrics.
  if (type == "DSM") { 
    gf <- data.frame(MEC = MEC, 
                     concordance = ccc, 
                     MSE = SEP2, 
                     RMSE = SEP, 
                     bias = bias, 
                     row.names = NULL)
  } else if (type == "spec") { 
    gf <- data.frame(MEC = MEC, 
                     concordance = ccc, 
                     MSE = SEP2, 
                     RMSE = SEP, 
                     bias = bias, 
                     MSEc = SEP2c, 
                     RMSEc = SEPc, 
                     RPD = RPD, 
                     RPIQ = RPIQ, 
                     row.names = NULL)
  } else { 
    stop("ERROR: Revise the type of output you require. Select from either DSM or spec")
  } 
  return(gf)
}

```

### Setting the folder path
This is the directory where the data files are stored.

```{r setup, include=FALSE}
dir.p = 'D:/project_lowcost_nir/database' # homeoffice/
```

### Importing more ASD data and testing without the transferred data

## Paulinenaue / Peatlands
Reading the csv file containing the ASD data for Paulinenaue / Peatlands.

```{r setup, include=FALSE}
asd.paul = read.csv(paste0(dir.p, "/raw_data/paulinenaue/db_soil_spec_epsg32633.csv"), header = T, sep = ",")[-1]

head(asd.paul[, 1:10], 5) 

colnames(asd.paul)[1:42]
asd.paul = asd.paul[, c(1, 11:12, 42:ncol(asd.paul))]
head(asd.paul[, c(1:10, ncol(asd.paul))], 5) # how it should look like

asd.paul = asd.paul[complete.cases(asd.paul), ] # 262 obs
colnames(asd.paul) = gsub("X", "", names(asd.paul))
```


Plotting the ASD spectra for Peatlands.
```{r setup, include=FALSE}
matplot(as.numeric(colnames(asd.paul)[4:ncol(asd.paul)]), 
        t(asd.paul[,4:ncol(asd.paul)]),
        xlim = c(350, 2500),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)
axis(1, at = seq(350, 2500, by = 50), labels = T, las = 2)
mtext('ASD Spectra: Peatlands')
grid(lwd=1, nx=50, ny=50)

matplot(as.numeric(colnames(asd.vnir.osc)[4:ncol(asd.vnir.osc)]), 
        t(asd.vnir.osc[,4:ncol(asd.vnir.osc)]),
        xlim = c(350, 2500),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 1.0, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2,
        add = T)
```

Merging datasets for preprocessing
The datasets are merged for preprocessing.

```{r setup, include=FALSE}
asd.paul[1:5, c(1:10, ncol(asd.paul))]
colnames(asd.paul)[1] = "id"
asd.paul = asd.paul[, c(1, 3, 2, 4:ncol(asd.paul))]
colnames(asd.paul)[2:3] = c("tn.perc", "tc.perc")
asd.paul$tn.perc = asd.paul$tn.perc/10
nrow(asd.paul)

asd.osc[1:5, c(1:10, ncol(asd.osc))]
nrow(asd.osc)

asd.chem.de = rbind(asd.paul, asd.osc)
asd.chem.de[1:5, c(1:10, ncol(asd.chem.de))]
```

### Splice Correction applicable for ASD
Splice correction is a preprocessing step in spectral analysis that corrects the abrupt changes in reflectance values at the junction of two detectors in a spectrometer. This is applicable for ASD (Analytical Spectral Devices) spectrometers which have two detectors, one for the Visible and Near-Infrared (VNIR) range and another for the Short-Wave Infrared (SWIR) range.

```{r setup, include=FALSE}
asd.chem.de[1:5, c(1:10, ncol(asd.chem.de))]

db.asd.paul.spl = asd.paul
colnames(db.asd.paul.spl)[c(1:4, ncol(db.asd.paul.spl))]
head(db.asd.paul.spl[, c(1:4, ncol(db.asd.paul.spl))])
rownames(db.asd.paul.spl) = db.asd.paul.spl$id

class(db.asd.paul.spl)
colnames(db.asd.paul.spl)[1:10]
wav.asd = as.numeric(colnames(db.asd.paul.spl[, 4:ncol(db.asd.paul.spl)]))
db.asd.paul.spl.m = as.matrix(db.asd.paul.spl[, 4:ncol(db.asd.paul.spl)]) # assembling matrix
class(db.asd.paul.spl.m) # checking class

db.asd.paul.spl.c = prospectr::spliceCorrection(db.asd.paul.spl.m, wav.asd, 
                                               splice = c(1000, 1800), interpol.bands = 10)
class(db.asd.paul.spl.c)
colnames(db.asd.paul.spl.c)
head(db.asd.paul.spl.c[1:5, 1:5])
```

## Resampling from 1nm to 5 nm 

Resampling is a preprocessing step that changes the resolution of the spectra. Here, the spectra are resampled from a resolution of 1 nm to 5 nm.

```{r}
db.asd.paul.spl.c[1:5, c(1, ncol(db.asd.paul.spl.c))]
db.asd.paul.r = prospectr::resample(db.asd.paul.spl.c, 
                                    wav.asd, 
                                    seq(435, 2415, 5), 
                                    interpol = "spline") # window size of 5 nm


class(db.asd.paul.r)
```
## Savitzky-Golay

Savitzky-Golay is a smoothing filter used to smooth the spectra. It uses a sliding window approach and fits a polynomial of a certain degree to the data within the window at each point.

```{r}
db.asd.paul.sg = prospectr::savitzkyGolay(db.asd.paul.r[, 1:ncol(db.asd.paul.r)],
                                         m = 0, # differentiation order (derivative)
                                         p = 2, # polynomial order
                                         w = 11 # window size
) 
class(db.asd.paul.sg)
head(db.asd.paul.sg[1:5, 1:5])
which(colnames(db.asd.paul.sg) == "950") 
which(colnames(db.asd.paul.sg) == "1650") 
db.asd.paul.sg = db.asd.paul.sg[, 99:239]
db.asd.paul.sg[1:5, c(1, ncol(db.asd.paul.sg))]
```

## SNV - Select the range before perform it!!!

Standard Normal Variate (SNV) is a normalization method used to correct for multiplicative scatter effects in the spectra.

```{r}
db.asd.paul.snv = prospectr::standardNormalVariate(db.asd.paul.sg)

### Transforming to DF 
db.asd.paul.d = as.data.frame(db.asd.paul.snv) # converting back to DF
db.asd.paul.d[1:20, c(1:5, ncol(db.asd.paul.d))]
db.asd.paul.d$id = rownames(db.asd.paul.d)
db.asd.paul.d = db.asd.paul.d %>%
  dplyr::select(c('id'), everything())
head(db.asd.paul.d[1:5, c(1:2, ncol(db.asd.paul.d))])
```
### filtering from 950 - 1650 nm

The spectra are filtered to select only the wavelengths from 950 nm to 1650 nm.

```{r}
#which(colnames(db.asd.paul.d) == "950") 
#which(colnames(db.asd.paul.d) == "1650") 
#colnames(db.asd.paul.d)[c(1,100, 240)]
#db.asd.paul.d = db.asd.paul.d[, c(1, 100:240)]
head(db.asd.paul.d[1:5, c(1:2, ncol(db.asd.paul.d))])
```
### Merging with soil data

The spectral data is merged with the soil data.

```{r}
#asd.chem.de
asd.paul[1:5, c(1:5, ncol(asd.paul))]
nrow(asd.paul) #262 obs
#head(asd.chem.de[1:5, c(1:4, ncol(asd.chem.de))])
head(db.asd.paul.d[1:5, c(1:2, ncol(db.asd.paul.d))])
db.soil.asd.nir = merge(asd.paul[, 1:3], db.asd.paul.d, by="id")
head(db.soil.asd.nir[1:5, c(1:5, ncol(db.soil.asd.nir))])

#jpeg(paste0(dir.p, '/results/figs/splice_asd_osc_vnir.jpeg'), width = 10, height = 10, units = 'in', res = 300)
matplot(as.numeric(colnames(db.soil.asd.nir)[4:ncol(db.soil.asd.nir)]), 
        t(db.soil.asd.nir[,4:ncol(db.soil.asd.nir)]),
        xlim = c(950, 1650),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)
mtext('ASD Spectra: SG-SNV (955 - 1645 nm)')
grid(lwd=1, nx=50, ny=50)
#dev.off()
```
## seeing where the target fits in 

The target spectra are plotted to see where they fit in the overall spectra.

```{r}
asd.target = read.csv(paste0(dir.p, "/tables/db_model/db_target_asd_spec_cal.csv"), header = T, sep = ",")[-1]
head(asd.target[1:5, c(1:3, ncol(asd.target))])
colnames(asd.target) = gsub("NIR.", "", names(asd.target))
#asd.target = asd.target[, c(1, 3:141)]

matplot(as.numeric(colnames(asd.target)[2:ncol(asd.target)]), 
        t(asd.target[,2:ncol(asd.target)]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)
```


### Adding local data to regional model
#### Oscar
The local data from Oscar is added to the regional model. The data is read from a CSV file and preprocessed for further analysis.

```{r}
db.soil = read.csv(paste0(dir.p, '/tables/db_model/db_soil_oscar.csv'))
db.soil
db.soil$id = sprintf("osc_%02d", db.soil$id)
db.soil = db.soil[1:102,]
db.soil = db.soil[!(db.soil$tn.perc == "#N/A" & db.soil$tc.perc == "#N/A"), ]
db.soil[1:5, 1:8]

head(asd.target[1:5, c(1:3, ncol(asd.target))])

asd.soil.nir.osc = merge(db.soil[, c(1, 5:6)], asd.target, by="id")
```

The spectral data is merged with the soil data.

```{r}
#which(colnames(asd.vnir.osc) == "435") 
#which(colnames(asd.vnir.osc) == "2415") 
asd.soil.nir.osc[1:5, c(1:10, ncol(asd.soil.nir.osc))]
nrow(asd.soil.nir.osc)
ncol(asd.soil.nir.osc)
db.soil.asd.nir[1:5, c(1:10, ncol(db.soil.asd.nir))]
nrow(db.soil.asd.nir)
ncol(db.soil.asd.nir)

asd.chem.de = rbind(db.soil.asd.nir, asd.soil.nir.osc)

write.csv(asd.chem.de, paste0(dir.p, "/tables/db_model/db_soil_germany_asd_nir_snv.csv"))
```

### Exporting the table ----
```{r}
write.csv(db.soil.asd.nir, paste0(dir.p, '/tables/db_soil_asd_peat_nir.csv'))
```

### OSSSL model: Testing ASD large library SNV NIR range (955 - 1645 nm)
The OSSSL model is tested using a large library of ASD spectra. The spectra are preprocessed using Standard Normal Variate (SNV) and filtered to select only the wavelengths from 955 nm to 1645 nm.

```{r}
str(db.soil.asd.nir[, 1:10])
#colnames(db.soil.asd.nir)[2] = "tc.perc"

ctrl = caret::trainControl(method = 'repeatedcv', number = 10, repeats = 10, search = 'random')
cores = detectCores()-2
#asd.chem.de
#asd.chem.de[1:5, c(1:10, ncol(asd.chem.de))]
class(asd.chem.de)
str(asd.chem.de)
asd.chem.de$tc.perc = as.numeric(asd.chem.de$tc.perc)
## Cubist
cores # it has the number of cores from parallel::detectCores()-2
cl = parallel::makeCluster(cores) # Create a cluster using all cores
doParallel::registerDoParallel(cl)

system.time(ossl.tc.m <- caret::train(x=asd.chem.de[ ,4:ncol(asd.chem.de)], #ncol(db.soil.asd.nir)
                                                 y=asd.chem.de$tc.perc,
                                                 method = "cubist", 
                                                 trControl = ctrl,
                                                 importance = T))  
stopCluster(cl)

print(ossl.tc.m)

ossl.tc.m.var = varImp(ossl.tc.m, scale = T)

#jpeg(paste0(dir.p, '/results/figs_cub/varimp/soc_SNV_isc_nir_Var_Importance_Cub.jpeg'), width = 10, height = 10, units = 'in', res = 300)
plot(ossl.tc.m.var, top = 20, main = "Original data")
#dev.off()
```

## Predicting
### ASD data
```{r}
db.soil.val.all = read.csv(paste0(dir.p, "/tables/db_soil_asd_nir_val.csv"), header = T, sep = ",")[-1] 
str(db.soil.val.all)

db.asd.nir.val = read.csv(paste0(dir.p, "/tables/db_model/db_target_asd_spec_val.csv"), header = T, sep = ",")[-1] 
str(db.asd.nir.val)
colnames(db.asd.nir.val) = gsub("NIR.", "", names(db.asd.nir.val))
db.asd.nir.val[1:5, c(1:5, ncol(db.asd.nir.val))]
db.soil.val.all$tc.asd = stats::predict(ossl.tc.m,
                                       db.asd.nir.val[, 2:ncol(db.asd.nir.val)]) # Predicting using external data / ncol(db.asd.nir.val)

db.soil.val.all$tc.asd = round(db.soil.val.all$tc.asd, 2)

asd.nir.val.metrics = goof(observed = db.soil.val.all$tc.perc, 
                           predicted = db.soil.val.all$tc.asd,
                           type = 'spec')


#jpeg(paste0(dir.p, '/Results/figs_cub/soc_SNV_isc_nir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.asd', 'tc.perc')])
plot(db.soil.val.all$tc.asd, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 30),
     ylim = c(0, 30),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
#text(1.8, 0.5, bquote(RMSE == .(round(asd.nir.val.metrics$RMSE, 2))))
#text(1.8, 0.4, bquote(MEC == .(round(asd.nir.val.metrics$MEC, 2))))
#text(1.8, 0.3, bquote(CCC == .(round(asd.nir.val.metrics$concordance, 2))))
#text(1.8, 0.2, bquote(bias == .(round(asd.nir.val.metrics$bias, 2))))
#text(1.8, 0.1, bquote(RPIQ == .(round(asd.nir.val.metrics$RPIQ, 2))))
#dev.off()
```

### ISC data
```{r}
db.isc.nir.val = read.csv(paste0(dir.p, "/tables/db_model/db_source_spec_subcorr_val.csv"), header = T, sep = ",")[-1] 
str(db.isc.nir.val)
colnames(db.isc.nir.val) = gsub("NIR.", "", names(db.isc.nir.val))
db.isc.nir.val[1:5, c(1:5, ncol(db.isc.nir.val))]

db.soil.val.all$tc.isc = stats::predict(ossl.tc.m,
                                        db.isc.nir.val[, 2:ncol(db.isc.nir.val)]) # Predicting using external data

db.soil.val.all$tc.isc = round(db.soil.val.all$tc.isc, 2)

isc.nir.val.metrics = goof(observed = db.soil.val.all$tc.perc, 
                           predicted = db.soil.val.all$tc.isc,
                           type = 'spec')


#jpeg(paste0(dir.p, '/Results/figs_cub/soc_SNV_isc_nir_Cub_Obs_Pred.jpeg'), width = 10, height = 10, units = 'in', res = 300)
par(mar = c(6, 6, 6, 6))
summary(db.soil.val.all[, c('tc.isc', 'tc.perc')])
plot(db.soil.val.all$tc.asd, db.soil.val.all$tc.perc,
     xlab = 'Predicted (%)', #expression(paste('Predicted ', '(', 'mg', ' kg'^'-1',')')),
     ylab = 'Observed (%)', #expression(paste('Observed ', '(', 'mg', ' kg'^'-1',')')),
     xlim = c(0, 10),
     ylim = c(0, 10),
     type = 'p',
     pch =  '🔅', # 16,
     bty = 'L',
     cex = 0.9,
     cex.lab= 1.1,
     cex.axis=1.2,
     col = rgb(red=0.0 , green = 0.0, blue = 0.0, alpha = .8),
     family = 'A')
abline(0, 1, col='black', lty=1, lwd=1)
abline(lm(db.soil.val.all$tc.asd ~ db.soil.val.all$tc.perc), col = 'red', lty=2, lwd=1)
#text(1.8, 0.5, bquote(RMSE == .(round(asd.nir.val.metrics$RMSE, 2))))
#text(1.8, 0.4, bquote(MEC == .(round(asd.nir.val.metrics$MEC, 2))))
#text(1.8, 0.3, bquote(CCC == .(round(asd.nir.val.metrics$concordance, 2))))
#text(1.8, 0.2, bquote(bias == .(round(asd.nir.val.metrics$bias, 2))))
#text(1.8, 0.1, bquote(RPIQ == .(round(asd.nir.val.metrics$RPIQ, 2))))
#dev.off()

matplot(as.numeric(colnames(asd.chem.de)[4:ncol(asd.chem.de)]), 
        t(asd.chem.de[,4:ncol(asd.chem.de)]),
        xlim = c(950, 1650),
        ylim = c(0,1),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        col = rgb(red = 0.5, green = 0.5, blue = 0.5, alpha = 0.5),
        xlab =  'nm', #expression(paste('Wavelength(cm'^'-1',')'))
        ylab = 'Reflectance',
        las=1, font=2)
axis(1, at = seq(950, 1650, by = 10), labels = T, las = 2)
mtext('ASD Spectra: SG-SNV (955 - 1645 nm)')
grid(lwd=1, nx=50, ny=50)
#dev.off()
```

## seeing where the source fits in 
```{r}
matplot(as.numeric(colnames(db.isc.nir.val)[2:ncol(db.isc.nir.val)]), 
        t(db.isc.nir.val[,2:ncol(db.isc.nir.val)]),
        #xlim = c(400, 2450),
        #ylim = c(0,0.6),
        xaxt = 'n',
        frame.plot = FALSE,
        type = 'l',
        lty = 1,
        lwd = 1,
        col = rgb(red = 0.13, green = 0.55, blue = 0.13, alpha = 0.5),
        las=1, font=2,
        add = T)
```

```{r}
# Saving R workspace ----
save.image(file = paste0(dir.p, '/r_code/modelling_all_free_asd_library.RData'))
```